---
title: "Hands-on Exercise 2a"
author: "Jeounghyun Kim"
date-modified: "last-modified" 
execute:
  echo: true
  eval: true
  warning: false
  freeze: true
---

# **5  2nd Order Spatial Point Patterns Analysis Methods**

## **5.1 Overview**

-   **Second-order SPPA** studies **interpoint relationships**—how the presence of one point affects others—across spatial scales. It goes beyond first-order intensity (overall density) to test for **clustering, dispersion, or randomness**.

-   **Goal of the exercise:** use  [spatstat](https://cran.r-project.org/web/packages/spatstat/) to investigate the spatial point process of **childcare centres in Singapore**.

-   **Key questions:**

    1.  Are childcare centres **consistent with CSR** (complete spatial randomness)?

    2.  If not, **where (and at what scales) do higher concentrations** occur?

## **5.2 The data**

To provide answers to the questions above, two data sets will be used. They are:

-   *Child Care Services* data from data.gov.sg, a point feature data providing both location and attribute information of childcare centres.

-   *Master Plan 2019 Subzone Boundary (No Sea)*, a polygon feature data providing information of URA 2019 Master Plan Planning Subzone boundary data.

Both data sets are available at [Singapore’s open data portal](https://r4gdsa.netlify.app/data.gov.sg). They are provided in kml and geojson format. Students are free to download their preferred data format.

## **5.3 Installing and Loading the R packages**

-   **sf** — Core toolkit for importing, managing, and processing **vector** geospatial data in R.

-   **spatstat** — Functions for **spatial point pattern analysis** (1st/2nd order) and **KDE** surface creation.

-   **tmap** — Makes **cartographic-quality maps** (static or interactive via Leaflet).

-   **tidyverse** — Integrated suite for **data wrangling, analysis, and visualization** with consistent grammar and data structures.

```{r}
pacman::p_load(sf, spatstat, tmap, tidyverse)
```

## **5.4 Data Import and Preparation**

-   [Importing and Wrangling Geospatial Data Sets](file:///C:/tskam/r4gdsa/_book/chap04.html#importing-and-wrangling-geospatial-data-sets)

-   [Geospatial Data wrangling](file:///C:/tskam/r4gdsa/_book/chap04.html#geospatial-data-wrangling) for preparing geospatial data in ppp object class required by spatstat package.

```{r}
mpsz_cl      <- read_rds("data/mpsz_cl.rds")          # cleaned subzones (sf, EPSG:3414)
childcare_sf <- read_rds("data/childcare_sf.rds")     # childcare points (sf, EPSG:3414)

# 2) Ensure same CRS
stopifnot(st_crs(mpsz_cl) == st_crs(childcare_sf))

# 3) Create study window (owin) from subzone polygon
#    Use the unioned boundary to define the analysis region
sg_owin <- as.owin(st_union(mpsz_cl))   # owin window for spatstat

# 4) Build ppp from childcare sf points
coords <- st_coordinates(childcare_sf)  # extract x/y in meters
childcare_ppp <- ppp(x = coords[,1],
                     y = coords[,2],
                     window = sg_owin,
                     checkdup = FALSE)

# 5) (Optional) Rescale to kilometers for bandwidth/KDE convenience
childcare_ppp_km <- rescale(childcare_ppp, s = 1000, unitname = "km")

# 6) Quick sanity checks
summary(childcare_ppp_km)  # should show marks=NULL, window=owin, unit=km
```

```{r}
# 1) Extract target planning areas (sf)
pg <- mpsz_cl %>% filter(PLN_AREA_N == "PUNGGOL")
tm <- mpsz_cl %>% filter(PLN_AREA_N == "TAMPINES")
ck <- mpsz_cl %>% filter(PLN_AREA_N == "CHOA CHU KANG")
jw <- mpsz_cl %>% filter(PLN_AREA_N == "JURONG WEST")

# 2) Convert sf polygons to spatstat windows (owin)
pg_owin <- as.owin(pg)
tm_owin <- as.owin(tm)
ck_owin <- as.owin(ck)
jw_owin <- as.owin(jw)

# 3) Subset the childcare ppp by each window (meters)
childcare_pg_ppp <- childcare_ppp[pg_owin]
childcare_tm_ppp <- childcare_ppp[tm_owin]
childcare_ck_ppp <- childcare_ppp[ck_owin]
childcare_jw_ppp <- childcare_ppp[jw_owin]

# 4) Rescale to kilometres (for readable bandwidth/units) 
childcare_pg_ppp.km <- rescale.ppp(childcare_pg_ppp, 1000, "km")
childcare_tm_ppp.km <- rescale.ppp(childcare_tm_ppp, 1000, "km")
childcare_ck_ppp.km <- rescale.ppp(childcare_ck_ppp, 1000, "km")
childcare_jw_ppp.km <- rescale.ppp(childcare_jw_ppp, 1000, "km")
```

```{r}
# 5) Helper: draw a panel (white window + black points)
panel_ppp <- function(ppp_obj, title) {
  plot(Window(ppp_obj), col = "white", border = "grey30",
       main = title, axes = FALSE)
  plot(unmark(ppp_obj), add = TRUE,
       pch = 1, cex = 0.4, col = "black", lwd = 0.7) 
}  

# 6) 2×2 plot on screen
par(mfrow = c(2, 2), mar = c(2, 2, 2, 1), cex.main = 1.15)
panel_ppp(childcare_pg_ppp.km, "Punggol")
panel_ppp(childcare_tm_ppp.km, "Tampines")
panel_ppp(childcare_ck_ppp.km, "Choa Chu Kang")
panel_ppp(childcare_jw_ppp.km, "Jurong West")
```

## **5.5 Second-order Spatial Point Patterns Analysis**

## **5.6 Analysing Spatial Point Process Using G-Function**

The G function measures the distribution of the distances from an arbitrary event to its nearest event. In this section, you will learn how to compute G-function estimation by using [*Gest()*](https://rdrr.io/cran/spatstat/man/Gest.html) of **spatstat** package. You will also learn how to perform monta carlo simulation test using [*envelope()*](https://rdrr.io/cran/spatstat/man/envelope.html) of **spatstat** package.

### **5.6.1 Choa Chu Kang planning area**

#### 5.6.1.1 Computing G-function estimation

```{r}
set.seed(1234)
```

The code chunk below is used to compute G-function using *Gest()* of **spatat** package.

```{r}
G_CK = Gest(childcare_ck_ppp, correction = "border")
plot(G_CK, xlim=c(0,500))
```

#### 5.6.1.2 Performing Complete Spatial Randomness Test

To confirm the observed spatial patterns above, a hypothesis test will be conducted. The hypothesis and test are as follows:

Ho = The distribution of childcare services at Choa Chu Kang are randomly distributed.

H1= The distribution of childcare services at Choa Chu Kang are not randomly distributed.

The null hypothesis will be rejected if p-value is smaller than alpha value of 0.001.

Monte Carlo test with G-fucntion

```{r}
G_CK.csr <- envelope(childcare_ck_ppp, Gest, nsim = 999)
```

```{r}
plot(G_CK.csr)
```

### **5.6.2 Tampines planning area**

#### 5.6.2.1 Computing G-function estimation

```{r}
G_tm = Gest(childcare_tm_ppp, correction = "best")
plot(G_tm)
```

#### 5.6.2.2 Performing Complete Spatial Randomness Test

To confirm the observed spatial patterns above, a hypothesis test will be conducted. The hypothesis and test are as follows:

Ho = The distribution of childcare services at Tampines are randomly distributed.

H1= The distribution of childcare services at Tampines are not randomly distributed.

The null hypothesis will be rejected is p-value is smaller than alpha value of 0.001.

The code chunk below is used to perform the hypothesis testing.

```{r}
G_tm.csr <- envelope(childcare_tm_ppp, Gest, correction = "all", nsim = 999)
```

```{r}
plot(G_tm.csr)
```

## **5.7 Analysing Spatial Point Process Using F-Function**

The F function estimates the empty space function F(r) or its hazard rate h(r) from a point pattern in a window of arbitrary shape. In this section, you will learn how to compute F-function estimation by using [*Fest()*](https://rdrr.io/cran/spatstat/man/Fest.html) of **spatstat** package. You will also learn how to perform monta carlo simulation test using [*envelope()*](https://rdrr.io/cran/spatstat/man/envelope.html) of **spatstat** package.

### **5.7.1 Choa Chu Kang planning area**

#### 5.7.1.1 Computing F-function estimation

The code chunk below is used to compute F-function using *Fest()* of **spatat** package.

```{r}
F_CK = Fest(childcare_ck_ppp)
plot(F_CK)
```

### **5.7.2 Performing Complete Spatial Randomness Test**

To confirm the observed spatial patterns above, a hypothesis test will be conducted. The hypothesis and test are as follows:

Ho = The distribution of childcare services at Choa Chu Kang are randomly distributed.

H1= The distribution of childcare services at Choa Chu Kang are not randomly distributed.

The null hypothesis will be rejected if p-value is smaller than alpha value of 0.001.

Monte Carlo test with F-fucntion

```{r}
F_CK.csr <- envelope(childcare_ck_ppp, Fest, nsim = 999)
```

```{r}
plot(F_CK.csr)
```

### **5.7.3 Tampines planning area**

#### 5.7.3.1 Computing F-function estimation

Monte Carlo test with F-fucntion

```{r}
F_tm = Fest(childcare_tm_ppp, correction = "best")
plot(F_tm)
```

#### 5.7.3.2 Performing Complete Spatial Randomness Test

To confirm the observed spatial patterns above, a hypothesis test will be conducted. The hypothesis and test are as follows:

Ho = The distribution of childcare services at Tampines are randomly distributed.

H1= The distribution of childcare services at Tampines are not randomly distributed.

The null hypothesis will be rejected is p-value is smaller than alpha value of 0.001.

The code chunk below is used to perform the hypothesis testing.

```{r}
F_tm.csr <- envelope(childcare_tm_ppp, Fest, correction = "all", nsim = 999)
```

```{r}
plot(F_tm.csr)
```

## **5.8 Analysing Spatial Point Process Using K-Function**

K-function measures the number of events found up to a given distance of any particular event. In this section, you will learn how to compute K-function estimates by using [*Kest()*](https://rdrr.io/cran/spatstat/man/Kest.html) of **spatstat** package. You will also learn how to perform monta carlo simulation test using *envelope()* of spatstat package.

### **5.8.1 Choa Chu Kang planning area**

#### 5.8.1.1 Computing K-fucntion estimate

```{r}
K_ck = Kest(childcare_ck_ppp, correction = "Ripley")
plot(K_ck, . -r ~ r, ylab= "K(d)-r", xlab = "d(m)")
```

#### 5.8.1.2 Performing Complete Spatial Randomness Test

To confirm the observed spatial patterns above, a hypothesis test will be conducted. The hypothesis and test are as follows:

Ho = The distribution of childcare services at Choa Chu Kang are randomly distributed.

H1= The distribution of childcare services at Choa Chu Kang are not randomly distributed.

The null hypothesis will be rejected if p-value is smaller than alpha value of 0.001.

The code chunk below is used to perform the hypothesis testing.

```{r}
K_ck.csr <- envelope(childcare_ck_ppp, Kest, nsim = 99, rank = 1, glocal=TRUE)
```

```{r}
plot(K_ck.csr, . - r ~ r, xlab="d", ylab="K(d)-r")
```

### **5.8.2 Tampines planning area**

#### 5.8.2.1 Computing K-fucntion estimation

```{r}
K_tm = Kest(childcare_tm_ppp, correction = "Ripley")
plot(K_tm, . -r ~ r, 
     ylab= "K(d)-r", xlab = "d(m)", 
     xlim=c(0,1000))
```

#### 5.8.2.2 Performing Complete Spatial Randomness Test

To confirm the observed spatial patterns above, a hypothesis test will be conducted. The hypothesis and test are as follows:

Ho = The distribution of childcare services at Tampines are randomly distributed.

H1= The distribution of childcare services at Tampines are not randomly distributed.

The null hypothesis will be rejected if p-value is smaller than alpha value of 0.001.

The code chunk below is used to perform the hypothesis testing.

```{r}
K_tm.csr <- envelope(childcare_tm_ppp, Kest, nsim = 99, rank = 1, glocal=TRUE)
```

```{r}
plot(K_tm.csr, . - r ~ r, 
     xlab="d", ylab="K(d)-r", xlim=c(0,500))
```

## **5.9 Analysing Spatial Point Process Using L-Function**

In this section, you will learn how to compute L-function estimation by using [*Lest()*](https://rdrr.io/cran/spatstat/man/Lest.html) of **spatstat** package. You will also learn how to perform monta carlo simulation test using *envelope()* of spatstat package.

### **5.9.1 Choa Chu Kang planning area**

#### 5.9.1.1 Computing L Fucntion estimation

```{r}
L_ck = Lest(childcare_ck_ppp, correction = "Ripley")
plot(L_ck, . -r ~ r, 
     ylab= "L(d)-r", xlab = "d(m)")
```

#### 5.9.1.2 Performing Complete Spatial Randomness Test

To confirm the observed spatial patterns above, a hypothesis test will be conducted. The hypothesis and test are as follows:

Ho = The distribution of childcare services at Choa Chu Kang are randomly distributed.

H1= The distribution of childcare services at Choa Chu Kang are not randomly distributed.

The null hypothesis will be rejected if p-value if smaller than alpha value of 0.001.

The code chunk below is used to perform the hypothesis testing.

```{r}
L_ck.csr <- envelope(childcare_ck_ppp, Lest, nsim = 99, rank = 1, glocal=TRUE)
```

```{r}
plot(L_ck.csr, . - r ~ r, xlab="d", ylab="L(d)-r")
```

### **5.9.2 Tampines planning area**

#### 5.9.2.1 Computing L-fucntion estimate

```{r}
L_tm = Lest(childcare_tm_ppp, correction = "Ripley")
plot(L_tm, . -r ~ r, 
     ylab= "L(d)-r", xlab = "d(m)", 
     xlim=c(0,1000))
```

#### 5.9.2.2 Performing Complete Spatial Randomness Test

To confirm the observed spatial patterns above, a hypothesis test will be conducted. The hypothesis and test are as follows:

Ho = The distribution of childcare services at Tampines are randomly distributed.

H1= The distribution of childcare services at Tampines are not randomly distributed.

The null hypothesis will be rejected if p-value is smaller than alpha value of 0.001.

The code chunk below will be used to perform the hypothesis testing.

```{r}
L_tm.csr <- envelope(childcare_tm_ppp, Lest, nsim = 99, rank = 1, glocal=TRUE)
```

Then, plot the model output by using the code chun below.

```{r}
plot(L_tm.csr, . - r ~ r, 
     xlab="d", ylab="L(d)-r", xlim=c(0,500))
```
