---
title: "Hands-on Exercise 7"
author: "Jeounghyun Kim"
date: "October 16, 2025"
date-modified: "last-modified"
execute:
  cache: true
  echo: true
  eval: true
  warning: false
  freeze: false
---

# **13  Calibrating Hedonic Pricing Model for Private Highrise Property with GWR Method**

## **13.1 Overview**

-   A spatial technique for building [hedonic pricing](https://www.investopedia.com/terms/h/hedonicpricing.asp) models that capture **local variations** in relationships between variables.

-   Unlike **OLS**, GWR considers **spatial non-stationarity**, showing how factors affect prices differently across locations.

-   **Application:** Condominium resale prices (2015) using **structural** and **locational** variables.

## **13.2 The Data**

Two data sets will be used in this model building exercise, they are:

-   URA Master Plan subzone boundary in shapefile format (i.e. *MP14_SUBZONE_WEB_PL*)

-   condo_resale_2015 in csv format (i.e. *condo_resale_2015.csv*)

## **13.3 Getting Started**

Before we get started, it is important for us to install the necessary R packages into R and launch these R packages into R environment.

The R packages needed for this exercise are as follows:

-   R package for building OLS and performing diagnostics tests

    -   [**olsrr**](https://olsrr.rsquaredacademy.com/)

-   R package for calibrating geographical weighted family of models

    -   [**GWmodel**](https://cran.r-project.org/web/packages/GWmodel/index.html)

-   R package for multivariate data visualisation and analysis

    -   [**corrplot**](https://cran.r-project.org/web/packages/corrplot/vignettes/corrplot-intro.html)

-   Spatial data handling

    -   **sf**

-   Attribute data handling

    -   **tidyverse**, especially **readr**, **ggplot2** and **dplyr**

-   Choropleth mapping

    -   **tmap**

The code chunks below installs and launches these R packages into R environment.

```{r}
pacman::p_load(olsrr, corrplot, ggpubr, 
               sf, sfdep, GWmodel, tmap, 
               tidyverse, gtsummary,
               performance, RColorBrewer)
```

## **13.4 A short note about GWmodel**

-   **GWmodel** provides a suite of **localized spatial analysis tools**, including:

    -   Geographically Weighted (GW) **summary statistics**,

    -   **GW principal components analysis**,

    -   **GW discriminant analysis**,

    -   and various **GW regression models** (both basic and robust forms).

-   It is mainly used as an **exploratory spatial tool**, where results (parameters or outputs) are **mapped** to visualize spatial variations.

-   Such visualization often **guides further, more advanced statistical analysis**.

## **13.5 Geospatial Data Wrangling**

### **13.5.1 Importing geospatial data**

The geospatial data used in this hands-on exercise is called MP14_SUBZONE_WEB_PL. It is in ESRI shapefile format. The shapefile consists of URA Master Plan 2014’s planning subzone boundaries. Polygon features are used to represent these geographic boundaries. The GIS data is in svy21 projected coordinates systems.

The code chunk below is used to import *MP_SUBZONE_WEB_PL* shapefile by using `st_read()` of **sf** packages.

```{r}
mpsz = st_read(dsn = "data/geospatial", 
               layer = "MP14_SUBZONE_WEB_PL")
```

The report above shows that the R object used to contain the imported MP14_SUBZONE_WEB_PL shapefile is called *mpsz* and it is a simple feature object. The geometry type is *multipolygon*. it is also important to note that mpsz simple feature object does not have EPSG information.

### **13.5.2 Updating CRS information**

The code chunk below updates the newly imported *mpsz* with the correct ESPG code (i.e. 3414)

```{r}
mpsz_svy21 <- st_transform(mpsz, 3414)
```

After transforming the projection metadata, you can varify the projection of the newly transformed *mpsz_svy21* by using `st_crs()` of **sf** package.

The code chunk below will be used to verify the newly transformed *mpsz_svy21*.

```{r}
st_crs(mpsz_svy21)
```

Notice that the EPSG: is indicated as *3414* now.

Next, you will reveal the extent of *mpsz_svy21* by using `st_bbox()` of sf package.

```{r}
st_bbox(mpsz_svy21)
```

The print above reports the extent of mpsz_svy21 layer by its lower and upper limits.

## **13.6 Aspatial Data Wrangling**

### **13.6.1 Importing the aspatial data**

The *condo_resale_2015* is in csv file format. The codes chunk below uses `read_csv()` function of **readr** package to import *condo_resale_2015* into R as a tibble data frame called *condo_resale*.

```{r}
condo_resale = read_csv(
  "data/aspatial/Condo_resale_2015.csv")
```

After importing the data file into R, it is important for us to examine if the data file has been imported correctly.

The codes chunks below uses `glimpse()` to display the data structure of will do the job.

```{r}
glimpse(condo_resale)
```

```{r}
head(condo_resale$LONGITUDE)
```

```{r}
head(condo_resale$LATITUDE)
```

The print above reveals that the values of *LONGITITUDE* and *LATITUDE* fields are in decimal degree. Most probably wgs84 geographic coordinate system is used.

Next, `summary()` of base R is used to display the summary statistics of *cond_resale*tibble data frame.

```{r}
summary(condo_resale)
```

### **13.6.2 Converting aspatial data frame into a sf object**

-   The `condo_resale` tibble is **aspatial** (no geographic information).

-   It is converted into a **simple feature (sf) object** using the `st_as_sf()` function from the **sf package**, allowing spatial analysis and mapping.

```{r}
condo_resale.sf <- st_as_sf(condo_resale,
                            coords = c("LONGITUDE", "LATITUDE"),
                            crs=4326) %>%
  st_transform(crs=3414)
```

::: callout-note
Notice that `st_transform()` of **sf** package is used to convert the coordinates from wgs84 (i.e. crs:4326) to svy21 (i.e. crs=3414).
:::

Next, `head()` is used to list the content of *condo_resale.sf* object.

```{r}
head(condo_resale.sf)
```

Notice that the output is in point feature data frame.

## **13.7 Exploratory Data Analysis (EDA)**

In the section, you will learn how to use statistical graphics functions of **ggplot2**package to perform EDA.

### **13.7.1 EDA using statistical graphics**

We can plot the distribution of *SELLING_PRICE* by using appropriate Exploratory Data Analysis (EDA) as shown in the code chunk below.

```{r}
ggplot(data=condo_resale.sf, 
       aes(x=`SELLING_PRICE`)) +
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue")
```

The figure above reveals a right skewed distribution. This means that more condominium units were transacted at relative lower prices.

Statistically, the skewed dsitribution can be normalised by using log transformation. The code chunk below is used to derive a new variable called *LOG_SELLING_PRICE* by using a log transformation on the variable *SELLING_PRICE*. It is performed using `mutate()` of **dplyr** package.

```{r}
condo_resale.sf <- condo_resale.sf %>%
  mutate(`LOG_SELLING_PRICE` = log(SELLING_PRICE))
```

Now, you can plot the *LOG_SELLING_PRICE* using the code chunk below.

```{r}
ggplot(data=condo_resale.sf, 
       aes(x=`LOG_SELLING_PRICE`)) +
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue")
```

Notice that the distribution is relatively less skewed after the transformation.

### **13.7.2 Multiple Histogram Plots distribution of variables**

In this section, you will learn how to draw a small multiple histograms (also known as trellis plot) by using `ggarrange()` of [**ggpubr**](https://cran.r-project.org/web/packages/ggpubr/index.html) package.

The code chunk below is used to create 12 histograms. Then, `ggarrange()` is used to organised these histogram into a 3 columns by 4 rows small multiple plot.

```{r}
AREA_SQM <- ggplot(data=condo_resale.sf, 
                   aes(x= `AREA_SQM`)) +
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue")

AGE <- ggplot(data=condo_resale.sf, 
              aes(x= `AGE`)) +
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue")

PROX_CBD <- ggplot(data=condo_resale.sf, 
                   aes(x= `PROX_CBD`)) +
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue")

PROX_CHILDCARE <- ggplot(data=condo_resale.sf, 
                         aes(x= `PROX_CHILDCARE`)) + 
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue")

PROX_ELDERLYCARE <- ggplot(data=condo_resale.sf, 
                           aes(x= `PROX_ELDERLYCARE`)) +
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue")

PROX_URA_GROWTH_AREA <- ggplot(data=condo_resale.sf, 
                               aes(x= `PROX_URA_GROWTH_AREA`)) +
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue")

PROX_HAWKER_MARKET <- ggplot(data=condo_resale.sf, 
                             aes(x= `PROX_HAWKER_MARKET`)) +
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue")

PROX_KINDERGARTEN <- ggplot(data=condo_resale.sf, 
                            aes(x= `PROX_KINDERGARTEN`)) +
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue")

PROX_MRT <- ggplot(data=condo_resale.sf, 
                   aes(x= `PROX_MRT`)) +
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue")

PROX_PARK <- ggplot(data=condo_resale.sf, 
                    aes(x= `PROX_PARK`)) +
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue")

PROX_PRIMARY_SCH <- ggplot(data=condo_resale.sf, 
                           aes(x= `PROX_PRIMARY_SCH`)) +
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue")

PROX_TOP_PRIMARY_SCH <- ggplot(data=condo_resale.sf, 
                               aes(x= `PROX_TOP_PRIMARY_SCH`)) +
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue")

ggarrange(AREA_SQM, AGE, PROX_CBD, PROX_CHILDCARE, 
          PROX_ELDERLYCARE, PROX_URA_GROWTH_AREA,
          PROX_HAWKER_MARKET, PROX_KINDERGARTEN, PROX_MRT,
          PROX_PARK, PROX_PRIMARY_SCH, PROX_TOP_PRIMARY_SCH,  
          ncol = 3, nrow = 4)
```

### **13.7.3 Drawing Statistical Point Map**

Lastly, we want to reveal the geospatial distribution condominium resale prices in Singapore. The map will be prepared by using **tmap** package.

First, we will turn on the interactive mode of tmap by using the code chunk below.

```{r}
tmap_mode("plot")
```

Next, the code chunks below is used to create an interactive point symbol map.

```{r}
tm_shape(mpsz_svy21)+
  tm_polygons() +
tm_shape(condo_resale.sf) +  
  tm_dots(fill = "SELLING_PRICE",
          fill_alpha = 0.6,
          size = 0.5,
          fill.scale = tm_scale_intervals(
            style="quantile")) +
  tm_view(set_zoom_limits = c(11,14))
```

Notice that [`tm_dots()`](https://www.rdocumentation.org/packages/tmap/versions/2.2/topics/tm_symbols) is used instead of `tm_bubbles()`.

`set_zoom_limits` argument of `tm_view()` sets the minimum and maximum zoom level to 11 and 14 respectively.

Before moving on to the next section, the code below will be used to turn R display into `plot` mode.

```{r}
tmap_mode("plot")
```

## **13.8 Hedonic Pricing Modelling in R**

In this section, you will learn how to building hedonic pricing models for condominium resale units using [`lm()`](https://www.rdocumentation.org/packages/stats/versions/3.5.2/topics/lm) of R base.

### **13.8.1 Simple Linear Regression Method**

First, we will build a simple linear regression model by using *SELLING_PRICE* as the dependent variable and *AREA_SQM* as the independent variable.

```{r}
condo.slr <- lm(formula=SELLING_PRICE ~ AREA_SQM, 
                data = condo_resale.sf)
```

`lm()` returns an object of class “lm” or for multiple responses of class c(“mlm”, “lm”).

The functions `summary()` and `anova()` can be used to obtain and print a summary and analysis of variance table of the results. The generic accessor functions coefficients, effects, fitted.values and residuals extract various useful features of the value returned by `lm`.

```{r}
summary(condo.slr)
```

-   The regression equation is **y = -258121.1 + 14719x₁**, meaning that **SELLING_PRICE** increases by about **14,719** for each unit increase in **AREA_SQM**.

-   **R² = 0.4518**, indicating the model explains **about 45% of the variance** in resale prices.

-   The **p-value (\< 0.0001)** shows the model is **statistically significant**, so we reject the null hypothesis that the mean alone predicts prices well.

-   Both **Intercept (B₀)** and **AREA_SQM (B₁)** have **p-values \< 0.001**, confirming they are **significant predictors**.

-   The **best-fit line** can be visualized in a scatterplot using **`geom_smooth(method = "lm")`** in **ggplot2**.

```{r}
ggplot(data=condo_resale.sf,  
       aes(x=`AREA_SQM`, y=`SELLING_PRICE`)) +
  geom_point() +
  geom_smooth(method = lm)
```

Figure above reveals that there are a few statistical outliers with relatively high selling prices.

### **13.8.2 Multiple Linear Regression Method**

#### 13.8.2.1 Visualising the relationships of the independent variables

-   Before building a **multiple regression model**, it is essential to confirm that **independent variables are not highly correlated**.

-   **High correlation** among predictors can **distort model accuracy** — a problem known as **multicollinearity**.

-   A **correlation matrix** helps visualize relationships between variables.

-   In this exercise, the **`corrplot`** package (and `pairs()` in R) is used to **plot a scatterplot matrix** showing correlations among independent variables in `condo_resale`.

```{r}
corrplot(cor(condo_resale[, 5:23]), 
         diag = FALSE, 
         order = "AOE",
         tl.pos = "td", 
         tl.cex = 0.5, 
         method = "number", 
         type = "upper")
```

Matrix reorder is very important for mining the hiden structure and patter in the matrix. There are four methods in corrplot (parameter order), named “AOE”, “FPC”, “hclust”, “alphabet”. In the code chunk above, AOE order is used. It orders the variables by using the *angular order of the eigenvectors* method suggested by [Michael Friendly](https://www.datavis.ca/papers/corrgram.pdf).

From the scatterplot matrix, it is clear that ***Freehold*** is highly correlated to ***LEASE_99YEAR***. In view of this, it is wiser to only include either one of them in the subsequent model building. As a result, ***LEASE_99YEAR*** is excluded in the subsequent model building.

### **13.8.3 Building a hedonic pricing model using multiple linear regression method**

The code chunk below using `lm()` to calibrate the multiple linear regression model.

```{r}
condo.mlr <- lm(
  formula = SELLING_PRICE ~ AREA_SQM + AGE  + 
    PROX_CBD + PROX_CHILDCARE + PROX_ELDERLYCARE +
    PROX_URA_GROWTH_AREA + PROX_HAWKER_MARKET   +
    PROX_KINDERGARTEN   + PROX_MRT  + PROX_PARK +
    PROX_PRIMARY_SCH + PROX_TOP_PRIMARY_SCH +
    PROX_SHOPPING_MALL  + PROX_SUPERMARKET + 
    PROX_BUS_STOP   + NO_Of_UNITS + 
    FAMILY_FRIENDLY + FREEHOLD, 
  data=condo_resale.sf)
summary(condo.mlr)
```

::: callout-note
The code chunk above consists of two parts: - `lm()` of Base R is used to calibrate a multiple linear regression model. The model output is stored in an lm object called *condo.mlr*. - `summary()` is used to print the model output.
:::

### **13.8.4 Revising the model**

With reference to the report above, it is clear that not all the independent variables are statistically significant. We will revised the model by removing those variables which are not statistically significant.

Now, we are ready to calibrate the revised model by using the code chunk below.

```{r}
condo.mlr1 <- lm(
  formula = SELLING_PRICE ~ AREA_SQM + AGE + 
    PROX_CBD + PROX_CHILDCARE + PROX_MRT +
    PROX_ELDERLYCARE    + PROX_URA_GROWTH_AREA +
    PROX_PARK   + PROX_PRIMARY_SCH + 
    PROX_SHOPPING_MALL  + PROX_BUS_STOP + 
    NO_Of_UNITS + FAMILY_FRIENDLY + FREEHOLD,
  data=condo_resale.sf)
summary(condo.mlr1)
```

The output above reveals that all explanatory variables are statistically significant at 95% confident level.

### **13.8.5 Preparing Publication Quality Table: gtsummary method**

The [**gtsummary**](https://www.danieldsjoberg.com/gtsummary/index.html) package provides an elegant and flexible way to create publication-ready summary tables in R.

In the code chunk below, [`tbl_regression()`](https://www.danieldsjoberg.com/gtsummary/reference/tbl_regression.html) is used to create a well formatted regression report.

```{r}
tbl_regression(condo.mlr1, intercept = TRUE)
```

With gtsummary package, model statistics can be included in the report by either appending them to the report table by using [`add_glance_table()`](https://www.danieldsjoberg.com/gtsummary/reference/add_glance.html) or adding as a table source note by using [`add_glance_source_note()`](https://www.danieldsjoberg.com/gtsummary/reference/add_glance.html) as shown in the code chunk below.

```{r}
tbl_regression(condo.mlr1, 
               intercept = TRUE) %>% 
  add_glance_source_note(
    label = list(sigma ~ "\U03C3"),
    include = c(r.squared, adj.r.squared, 
                AIC, statistic,
                p.value, sigma))
```

For more customisation options, refer to [Tutorial: tbl_regression](https://www.danieldsjoberg.com/gtsummary/articles/tbl_regression.html)

### **13.8.6 Regression Diagnostics**

-   **Regression diagnostics** are procedures to verify if a regression model meets key assumptions and fits the data properly.

-   They help detect issues like **non-linearity**, **non-normal errors**, **heteroskedasticity**, and **influential observations** to ensure reliable results.

-   The **`olsrr`** R package provides tools for performing OLS regression diagnostics, including:

    -   comprehensive regression output

    -   residual diagnostics

    -   measures of influence

    -   heteroskedasticity tests

    -   collinearity diagnostics

    -   model fit assessment

    -   variable contribution assessment

    -   variable selection procedures

#### 13.8.6.1 Multicollinearity test

-   **Multicollinearity** occurs when independent variables are **highly correlated**, meaning changes in one variable are linked to changes in another.

-   This makes it difficult for the model to **separate each variable’s effect** on the dependent variable.

-   High multicollinearity causes **unstable and unreliable coefficients**, which can lead to **misleading interpretations**.

-   To detect it, the  [`check_collinearity()`](https://easystats.github.io/performance/reference/check_collinearity.html) function from the **`performance`** package is used.

```{r}
check_collinearity(condo.mlr1)
```

```{r}
mlr.vif <- check_collinearity(condo.mlr1)
plot(mlr.vif)
```

Since the VIF of the independent variables are less than 10. We can safely conclude that there are no sign of multicollinearity among the independent variables.

#### 13.8.6.2 Test for Non-Linearity

In multiple linear regression, it is important for us to test the assumption that linearity and additivity of the relationship between dependent and independent variables.

In the code chunk below, the [`ols_plot_resid_fit()`](https://olsrr.rsquaredacademy.com/reference/ols_plot_resid_fit.html) of **olsrr** package is used to perform linearity assumption test.

```{r}
check_model(condo.mlr1,
            check = "linearity")
```

The figure above reveals that most of the data poitns are scattered around the 0 line, hence we can safely conclude that the relationships between the dependent variable and independent variables are linear.

#### 13.8.6.3 Test for Normality Assumption

The normality assumption test for multiple linear regression checks **if the model’s residuals (the differences between observed and predicted values) are normally distributed**. This is crucial for accurate hypothesis testing and confidence intervals. To test this, you can use visual methods like histograms and Q-Q plots of the residuals, or conduct statistical tests like Shapiro-Wilk test and Kolmogorov-Smirnov test.

In the code chunk below, `check_normality()` of performance package is used to perform normality assumption test on condo.mlr1 model.

```{r}
check_normality(condo.mlr1)
```

The print above reveals that the p-value of the normality assumption test is less than alpha value of 0.05. Hence we reject the normality assumption at 95% confident level.

-   ::: callout-note
    -   `check_normality()` calls *stats::shapiro.test* and checks the standardized residuals (or studentized residuals for mixed models) for normal distribution.

    -   Note that this formal test almost always yields significant results for the distribution of residuals and visual inspection (e.g. Q-Q plots) are preferable.
    :::

Instead of showing the test statistic, `plot()` of **see** package can be used to plot a the output of `check_normality()` for visual inspection as shown below.

```{r}
plot(check_normality(condo.mlr1), 
     type = "qq")
```

Q-Q plot above below shows that majority of the data points are felt along the zero line.

Another way to check for normality assumption visual is by using [`check_model()`](https://easystats.github.io/performance/reference/check_model.html)of performance package as shown in the code chunk below.

```{r}
check_model(condo.mlr1,
            check= "normality")
```

The figure reveals that the residual of the multiple linear regression model (i.e. condo.mlr1) is resemble normal distribution.

#### 13.8.6.4 Testing for Spatial Autocorrelation

The hedonic model we try to build are using geographically referenced data. Hence, it is crucial to check for spatial autocorrelation because its presence can produce unreliable and misleading results. Traditional regression models, such as ordinary least squares (OLS), assume that observations are independent of one another. However, spatial data often violates this assumption.

Spatial autocorrelation is the correlation of a variable with itself across different spatial locations. Positive spatial autocorrelation means nearby features tend to be more similar, while negative autocorrelation means they tend to be more dissimilar. This phenomenon is based on the first law of geography: “Everything is related to everything else, but nearby things are more related than distant things”.

Ignoring spatial autocorrelation in a regression model can lead to serious statistical issues:

-   Biased and inefficient coefficient estimates: If autocorrelation is present, standard errors of the model coefficients can be wrong, leading to unreliable hypothesis tests (p-values). The model might appear more significant than it is.

-   Misleading significance tests: Standard regression models cannot distinguish between true explanatory power and the influence of spatial patterns, resulting in inaccurate p-values.

-   Model misspecification: Significant spatial autocorrelation in the regression residuals often signals that important explanatory variables are missing from the model. The spatial patterning of the residuals (over- and under-predictions) can provide clues about what these missing variables might be.

-   Inflated Type I error rates: Researchers might incorrectly reject a true null.

To test for spatial autocorrelation, We can run a Moran’s I test on the model’s residuals. Significant spatial autocorrelation in the residuals means the model is not capturing the full spatial story.

In order to perform spatial autocorrelation test, we need to export the residual of the hedonic pricing model and save it as a data frame first.

```{r}
mlr.output <- as.data.frame(condo.mlr1$residuals)
```

Next, we will join the newly created data frame with *condo_resale.sf* object.

```{r}
condo_resale.sf <- cbind(condo_resale.sf, 
                        condo.mlr1$residuals) %>%
rename(`MLR_RES` = `condo.mlr1.residuals`)
```

Next, we will use **tmap** package to display the distribution of the residuals on an interactive map.

The code churn below will turn on the interactive mode of tmap.

```{r}
tmap_mode("plot")
```

The code chunks below is used to create an interactive point symbol map.

```{r}
tm_shape(mpsz_svy21) +
  tm_polygons(fill_alpha = 0.4) +
tm_shape(condo_resale.sf) +  
  tm_dots(
    fill = "MLR_RES",
    size = 0.7,
    col = "black",
    fill.scale = tm_scale(
      n = 10,
      values = rev(
        brewer.pal(11, "RdBu")),
      style = "quantile",
      midpoint = NA),
    fill.legend = tm_legend(title = "Residuals")
  ) + 
  tm_title("LM Residuals (Quantile Classification)") +
  tm_layout(legend.outside = TRUE) +
  tm_view(set_zoom_limits = c(11,14))
```

Remember to switch back to “plot” mode before continue.

```{r}
tmap_mode("plot")
```

The figure above reveal that there is sign of spatial autocorrelation.

To proof that our observation is indeed true, Global Moran’s I test will be performed

First, we will compute the distance-based weight matrix by using [`st_dist_band()`](https://sfdep.josiahparry.com/reference/st_dist_band)function of **sfdep**.

```{r}
condo_resale.sf <- condo_resale.sf %>% 
  mutate(
    nb = st_dist_band(
      st_geometry(geometry), 
                 upper = 1500),
    wt = st_weights(
      nb, style = "W"),
    .before = 1) 
```

Next, [`global_moran_perm()`](https://sfdep.josiahparry.com/reference/global_moran_perm) of **sfdep** package will be used to perform Moran’s I test for residual spatial autocorrelation

```{r}
set.seed(1234)
global_moran_perm(
  condo_resale.sf$MLR_RES, 
  nb = condo_resale.sf$nb, 
  wt = condo_resale.sf$wt,
  alternative = "two.sided", 
  nsim = 499)
```

The Global Moran’s I test for residual spatial autocorrelation shows that it’s p-value is less than 0.00000000000000022 which is less than the alpha value of 0.05. Hence, we will reject the null hypothesis that the residuals are randomly distributed.

Since the Observed Global Moran I = 0.14389 which is greater than 0, we can infer than the residuals resemble cluster distribution.

## **13.9 Building Hedonic Pricing Models using GWmodel**

In this section, you are going to learn how to modelling hedonic pricing using both the fixed and adaptive bandwidth schemes

### **13.9.1 Building Fixed Bandwidth GWR Model**

#### 13.9.1.1 Computing fixed bandwith

In the code chunk below `bw.gwr()` of GWModel package is used to determine the optimal fixed bandwidth to use in the model. Notice that the argument ***adaptive*** is set to **FALSE** indicates that we are interested to compute the fixed bandwidth.

There are two possible approaches can be uused to determine the stopping rule, they are: CV cross-validation approach and AIC corrected (AICc) approach. We define the stopping rule using ***approach*** argeement.

```{r}
bw.fixed <- bw.gwr(
  formula = SELLING_PRICE ~ AREA_SQM + AGE  + 
    PROX_CBD + PROX_CHILDCARE + PROX_ELDERLYCARE    +
    PROX_URA_GROWTH_AREA + PROX_MRT + PROX_PARK + 
    PROX_PRIMARY_SCH + PROX_SHOPPING_MALL   + 
    PROX_BUS_STOP   + NO_Of_UNITS + FAMILY_FRIENDLY + 
    FREEHOLD, 
  data=condo_resale.sf, 
  approach="CV", 
  kernel="gaussian", 
  adaptive=FALSE, 
  longlat=FALSE)
```

The result shows that the recommended bandwidth is 971.3405 metres.

::: callout-note
(Quiz: Do you know why it is in metre?)

The **bandwidth** value is expressed in **meters** because it follows the **coordinate system unit** of the spatial dataset.\
In this case, the dataset `condo_resale.sf` uses a **projected coordinate system** (such as *SVY21 / EPSG:3414*, used in Singapore), where distances are measured in meters.\
Also, since the argument `longlat = FALSE` was specified in the `bw.gwr()` function, the model assumes a **planar (Euclidean) distance**, not geographic degrees.

If `longlat = TRUE` were used instead, the data would be treated as **longitude–latitude coordinates**, and the bandwidth would be measured in **degrees** instead of meters.
:::

#### 13.9.1.2 GWModel method - fixed bandwith

Now we can use the code chunk below to calibrate the gwr model using fixed bandwidth and gaussian kernel.

```{r}
gwr.fixed <- gwr.basic(
  formula = SELLING_PRICE ~ AREA_SQM + AGE  + 
    PROX_CBD + PROX_CHILDCARE + PROX_ELDERLYCARE    +
    PROX_URA_GROWTH_AREA + PROX_MRT + PROX_PARK +
    PROX_PRIMARY_SCH + PROX_SHOPPING_MALL   + 
    PROX_BUS_STOP   + NO_Of_UNITS + FAMILY_FRIENDLY + 
    FREEHOLD, 
  data=condo_resale.sf, 
  bw=bw.fixed, 
  kernel = 'gaussian', 
  longlat = FALSE)
```

The output is saved in a list of class “gwrm”. The code below can be used to display the model output.

```{r}
gwr.fixed
```

The report shows that the AICc of the gwr is 42263.61 which is significantly smaller than the globel multiple linear regression model of 42967.1.

### **13.9.2 Building Adaptive Bandwidth GWR Model**

In this section, we will calibrate the gwr-based hedonic pricing model by using adaptive bandwidth approach.

#### 13.9.2.1 Computing the adaptive bandwidth

Similar to the earlier section, we will first use `bw.gwr()` to determine the recommended data point to use.

The code chunk used look very similar to the one used to compute the fixed bandwidth except the `adaptive` argument has changed to **TRUE**.

```{r}
bw.adaptive <- bw.gwr(
  formula = SELLING_PRICE ~ AREA_SQM + AGE  + 
    PROX_CBD + PROX_CHILDCARE + PROX_ELDERLYCARE    +
    PROX_URA_GROWTH_AREA + PROX_MRT + PROX_PARK +
    PROX_PRIMARY_SCH + PROX_SHOPPING_MALL   + 
    PROX_BUS_STOP   + NO_Of_UNITS + 
    FAMILY_FRIENDLY + FREEHOLD, 
  data=condo_resale.sf, 
  approach="CV", 
  kernel="gaussian", 
  adaptive=TRUE, 
  longlat=FALSE)
```

The result shows that the 30 is the recommended data points to be used.

#### 13.9.2.2 Constructing the adaptive bandwidth gwr model

Now, we can go ahead to calibrate the gwr-based hedonic pricing model by using adaptive bandwidth and gaussian kernel as shown in the code chunk below.

```{r}
gwr.adaptive <- gwr.basic(
  formula = SELLING_PRICE ~ AREA_SQM + AGE + 
    PROX_CBD + PROX_CHILDCARE + PROX_ELDERLYCARE +
    PROX_URA_GROWTH_AREA + PROX_MRT + PROX_PARK +
    PROX_PRIMARY_SCH + PROX_SHOPPING_MALL + PROX_BUS_STOP +
    NO_Of_UNITS + FAMILY_FRIENDLY + FREEHOLD,
  data=condo_resale.sf, 
  bw=bw.adaptive, 
  kernel = 'gaussian', 
  adaptive=TRUE, 
  longlat = FALSE)
```

The code below can be used to display the model output.

```{r}
gwr.adaptive
```

The report shows that the AICc the adaptive distance gwr is 41982.22 which is even smaller than the AICc of the fixed distance gwr of 42263.61.

### **13.9.3 Visualising GWR Output**

-   **Condition Number:** checks local multicollinearity; values \>30 indicate instability.

-   **Local R²:** measures local model fit (0–1 range); low values suggest poor local performance.

-   **Predicted Values:** fitted y-values estimated by GWR.

-   **Residuals:** difference between observed and predicted values; standardized residuals (mean = 0, SD = 1) can be mapped to visualize spatial error patterns.

-   **Coefficient Standard Errors:** indicate reliability of coefficient estimates; large errors may signal local collinearity.

-   All results are stored in a **SpatialPointsDataFrame/SpatialPolygonsDataFrame (SDF)**, including fitted values, coefficients, standard errors, and t-values.

### **13.9.4 Converting SDF into *sf* data.frame**

To visualise the fields in **SDF**, we need to first covert it into **sf** data.frame by using the code chunk below.

```{r}
condo_resale.sf.adaptive <-
  st_as_sf(gwr.adaptive$SDF) %>%
  st_transform(crs=3414)
```

Next, `glimpse()` is used to display the content of *condo_resale.sf.adaptive* sf data frame.

```{r}
glimpse(condo_resale.sf.adaptive)
```

```{r}
summary(gwr.adaptive$SDF$yhat)
```

### **13.9.5 Visualising local R2**

The code chunks below is used to create an interactive point symbol map.

```{r}
tmap_mode("plot")
tm_shape(mpsz_svy21)+
  tm_polygons(fill_alpha = 0.1) +
tm_shape(condo_resale.sf.adaptive) +  
  tm_dots(col = "Local_R2",
          border.col = "gray60",
          border.lwd = 1) +
  tm_view(set.zoom.limits = c(11,14))

tmap_mode("plot")
```

### **13.9.6 Visualising coefficient estimates**

The code chunks below is used to create an interactive point symbol map.

```{r}
tmap_mode("plot")
AREA_SQM_SE <- tm_shape(mpsz_svy21)+
  tm_polygons(alpha = 0.1) +
tm_shape(condo_resale.sf.adaptive) +  
  tm_dots(col = "AREA_SQM_SE",
          border.col = "gray60",
          border.lwd = 1) +
  tm_view(set.zoom.limits = c(11,14))

AREA_SQM_TV <- tm_shape(mpsz_svy21)+
  tm_polygons(alpha = 0.1) +
tm_shape(condo_resale.sf.adaptive) +  
  tm_dots(col = "AREA_SQM_TV",
          border.col = "gray60",
          border.lwd = 1) +
  tm_view(set.zoom.limits = c(11,14))

tmap_arrange(AREA_SQM_SE, AREA_SQM_TV, 
             asp=1, ncol=2,
             sync = TRUE)

tmap_mode("plot")
```

## **13.10 Reference**

Gollini I, Lu B, Charlton M, Brunsdon C, Harris P (2015) “GWmodel: an R Package for exploring Spatial Heterogeneity using Geographically Weighted Models”. *Journal of Statistical Software*, 63(17):1-50, http://www.jstatsoft.org/v63/i17/

Lu B, Harris P, Charlton M, Brunsdon C (2014) “The GWmodel R Package: further topics for exploring Spatial Heterogeneity using GeographicallyWeighted Models”. *Geo-spatial Information Science* 17(2): 85-101, http://www.tandfonline.com/doi/abs/10.1080/1009502.2014.917453
