---
title: "In-class Exercise 8"
author: "Jeounghyun Kim"
date: "Oct 25, 2025"
date-modified: "last-modified" 
execute:
  echo: true
  eval: true
  warning: false
  freeze: true
---

```{r}
pacman::p_load(httr,tidyverse, sf, tmap, jsonlite, progress)
```

# Geocoding HDB resale data

## Basic version

```{r}
HDBresale <- read_csv(
  "data/rawdata/HDBresale.csv") %>%
  filter(flat_type == "4 ROOM") %>%
  filter(month >= "2025-09" & month <"2025-10")
```

```{r}
HDBresale$street_name <- gsub(
  "ST\\.", "SAINT",
  HDBresale$street_name
)
```

```{r}
geocode <- function(block, streetname) {
  base_url <- "https://onemap.gov.sg/api/common/elastic/search"
  address <- paste(block, streetname, sep = " ")
  query <- list(
    "searchVal" = address,
    "returnGeom" = "Y",
    "getAddrDetails" = "N",
    "pageNum" = "1"
  )

  res <- GET(base_url, query = query)
  restext <- content(res, as = "text")

  output <- fromJSON(restext) %>%
    as.data.frame() %>%
    select(results.LATITUDE, results.LONGITUDE)

  return(output)
}
```

```{r}
HDBresale$LATITUDE <- 0
HDBresale$LONGITUDE <- 0

for (i in 1:nrow(HDBresale)) {
  temp_output <- geocode(HDBresale[i, 4], HDBresale[i, 5])
  
  HDBresale$LATITUDE[i] <- temp_output$results.LATITUDE
  HDBresale$LONGITUDE[i] <- temp_output$results.LONGITUDE
}

```

```{r}
HDBresale_sf <- HDBresale %>%
  st_as_sf(
    coords = c("LONGITUDE", "LATITUDE"),
    crs = 4326
  ) %>%
  st_transform(crs = 3414)
```

```{r}
preschool <- st_read("data/rawdata/PreSchoolsLocation.kml") %>%
  st_transform(crs = 3414)
```

```{r}
within_350m <- st_is_within_distance(
  HDBresale_sf, preschool, dist = 350)

HDBresale_sf <- HDBresale_sf %>%
  mutate(preschool_350m = map_int(within_350m, length))
```

# Predicting

```{r}
pacman::p_load(
  sf, spdep, GWmodel, SpatialML,
  tmap, rsample, Metrics, tidyverse,
  knitr, kableExtra, spatialRF,
  randomForestExplainer
)
```

## Preparing Data

```{r}
mdata <- read_rds("data/rawdata/mdata.rds")
```

### Data Sampling

```{r}
set.seed(1234)
HDB_sample <-mdata %>%
  sample_n(1500)
```

### Checking of overlapping point

When using GWmodel to calibrate explanatory or predictive models, it is very important to ensure that there are no overlapping point features.

The code chunk below is used to check if there are overlapping point features.

```{r}
overlapping_points <- HDB_sample %>%
  mutate(overlap = lengths(st_equals(., .)) > 1)
```

### Spatial Jitter

```{r}
HDB_sample <- HDB_sample %>%
  st_jitter(amount=1)
```

Check again for overlapping points

```{r}
overlapping_points <- HDB_sample %>%
  mutate(overlap = lengths(st_equals(., .)) > 1)
```

we realized no overlapped.

```{r}
set.seed(1234)

resale_split <- initial_split(HDB_sample, 
                              prop = 6.67/10)

train_data <- training(resale_split)
test_data  <- testing(resale_split)
```

```{r}
price_mlr <- lm(resale_price ~ floor_area_sqm +
                  storey_order + remaining_lease_mths +
                  PROX_CBD + PROX_ELDERLYCARE + PROX_HAWKER +
                  PROX_MRT + PROX_PARK + PROX_MALL + 
                  PROX_SUPERMARKET + WITHIN_350M_KINDERGARTEN +
                  WITHIN_350M_CHILDCARE + WITHIN_350M_BUS +
                  WITHIN_1KM_PRISCH,
                data=train_data)
olsrr::ols_regress(price_mlr)
```

```{r}
bw_adaptive <- bw.gwr(resale_price ~ floor_area_sqm +
                  storey_order + remaining_lease_mths +
                  PROX_CBD + PROX_ELDERLYCARE + PROX_HAWKER +
                  PROX_MRT + PROX_PARK + PROX_MALL + 
                  PROX_SUPERMARKET + WITHIN_350M_KINDERGARTEN +
                  WITHIN_350M_CHILDCARE + WITHIN_350M_BUS +
                  WITHIN_1KM_PRISCH,
                  data=train_data,
                  approach="CV",
                  kernel="gaussian",
                  adaptive=TRUE,
                  longlat=FALSE)
```

```{r}
write_rds(bw_adaptive, "data/model/bw_adaptive.rds")
```

```{r}
bw_adaptive
```

```{r}
gwr_adaptive <-read_rds("data/model/bw_adaptive.rds")
```
