---
title: "Hands-on Exercise 3"
author: "Jeounghyun Kim"
date: 2025-09-09
date-format: "MMMM D, YYYY"
execute:
  echo: true
  eval: true
  warning: false
  freeze: true
---

# **6  Spatio-Temporal Point Patterns Analysis**

## 6.1 Overview

**Definition**\
A spatio-temporal point process is a random collection of points where each point represents both the location and time of an event. It is also referred to as a space-time or spatial-temporal point process.

**Examples of Events**

-   Disease incidence

-   Sightings or births of a species

-   Natural hazard events such as fires, earthquakes, lightning strikes, tsunamis, volcanic eruptions

**Why It Matters**

-   The amount of geographically and temporally indexed data is rapidly growing across many fields, making spatio-temporal analysis increasingly important.

-   In the past decade, several spatio-temporal analysis methods have been developed and implemented in R.

**Focus of Lesson 3**

-   Demonstrate how different R packages can be combined to perform spatio-temporal point pattern analysis in a guided and intuitive way.

-   Case study: Forest fire events in Kepulauan Bangka Belitung, Indonesia

-   Period: January 1, 2023 – December 31, 2023

-   Purpose: To illustrate methods, procedures, and interpretations.

## **6.2 Learning Outcome**

### **6.2.1 The research questions**

The specific questions we would like to answer are:

-   Are the forest fire locations in Kepulauan Bangka Belitung spatially and spatio-temporally independent?

<!-- -->

-   If not, where and when do the observed forest fire locations tend to cluster?

## **6.3 The data**

**Data Sets Used**

1.  **forestfires (CSV file)**

    -   Contains forest fire locations detected by the MODIS sensor.

    -   Data source: Fire Information for Resource Management System (FIRMS).

    -   For this exercise: only forest fires within **Kepulauan Bangka Belitung** are used.

2.  **Kepulauan_Bangka_Belitung (ESRI shapefile)**

    -   Shows sub-district (kelurahan) boundaries of Kepulauan Bangka Belitung.

    -   Data source: Indonesia Geospatial Portal.

    -   Original coverage: entire Indonesia.

    -   For this exercise: only sub-districts within **Kepulauan Bangka Belitung** are extracted.

## **6.4 Installing and Loading the R packages**

For the purpose of this study, six R packages will be used. They are:

-   [**sf**](https://r-spatial.github.io/sf/) provides functions for importing processing and wrangling geospatial data,,

-   [**raster**](https://rspatial.org/raster/) for handling raster data in R,

-   [**spatstat**](https://cran.r-project.org/web/packages/spatstat/index.html) for performing Spatial Point Patterns Analysis such as kcross, Lcross, etc.,

-   [**sparr**](https://tilmandavies.github.io/sparr/index.html) provides functions to estimate fixed and adaptive kernel-smoothed spatial relative risk surfaces via the density-ratio method and perform subsequent inference. Fixed-bandwidth spatiotemporal density and relative risk estimation is also supported

-   [**tmap**](https://r-tmap.github.io/tmap/) provides functions to produce cartographic quality thematic maps, and

-   [**tidyverse**](https://www.tidyverse.org/), a family of R packages that provide functions to perform common data science tasks including and not limited to data import, data transformation, data wrangling and data visualisation.

::: panel-tabset
### DIY

Using the steps you learned from previous chapter, write a code chunk to load the packages above onto R environment.

### The solution

```{r}
pacman::p_load(sf, terra, spatstat, sparr, tmap, tidyverse)
```
:::

## **6.5 Importing and Preparing Study Area**

### **6.5.1 Importing study area**

Code chunk below is used import study area (i.e. Kepulauan Bangka Belitung) into R environment.

```{r}
kbb <- st_read(dsn = "data/rawdata",
               layer = "Kepulauan_Bangka_Belitung")
```

The revised code chunk.

```{r}
kbb_sf <- st_read(dsn="data/rawdata",
               layer = "Kepulauan_Bangka_Belitung") %>%
  st_union() %>%
  st_zm(drop = TRUE, what = "ZM") %>%
  st_transform(crs = 32748)
```

### **6.5.2 Converting OWIN**

Next, `as.owin()` is used to convert *kbb* into an **owin** object.

```{r}
kbb_owin <- as.owin(kbb_sf)
kbb_owin
```

Next, `class()` is used to confirm if the output is indeed an owin object.

```{r}
class(kbb_owin)
```

## **6.6 Importing and Preparing Forest Fire data**

```{r}
fire_sf <- read_csv("data/rawdata/forestfires.csv") %>%
  st_as_sf(coords = c("longitude", "latitude"),
                       crs = 4326) %>%
  st_transform(crs = 32748)
```

Because **ppp** object only accept numerical or character as mark. The code chunk below is used to convert data type of *acq_date* to numeric.

```{r}
fire_sf <- fire_sf %>% 
  mutate(DayofYear = yday(acq_date)) %>%
  mutate(Month_num = month(acq_date)) %>%
  mutate(Month_fac = month(acq_date, 
                           label = TRUE, 
                           abbr = FALSE))
```

## **6.7 Visualising the Fire Points**

### **6.7.1 Overall plot**

::: panel-tabset
### DIY

Using the steps you learned in Hands-on Exercise 2, prepare a point symbol map showing the distribution of fire points. The map should look similar to the figure below.

### The solution

```{r}
tm_shape(kbb_sf)+ tm_polygons() + tm_shape(fire_sf) + tm_dots()
```
:::

### **6.7.2 Visuaising geographic distribution of forest fires by month**

::: panel-tabset
### DIY

Using the steps you learned in Hands-on Exercise 2, prepare a point symbol map showing the monthly geographic distribution of forest fires in 2023. The map should look similar to the figure below.

### The solution

```{r}
tm_shape(kbb_sf)+
  tm_polygons() +
tm_shape(fire_sf) +
  tm_dots(size = 0.1) +
tm_facets(by="Month_fac", 
            free.coords=FALSE, 
            drop.units = TRUE)
```
:::

## **6.8 Computing STKDE by Month**

In this section, you will learn how to compute STKDE by using [`spattemp.density()`](https://tilmandavies.github.io/sparr/reference/spattemp.density.html) of **sparr** package. Before using the function, it is highly recommended you read the function’s reference guide in detail in order to understand the input data requirements and the output object generated.

### **6.8.1 Extracting forest fires by month**

The code chunk below is used to remove the unwanted fields from *fire_sf* sf data.frame. This is because `as.ppp()` only need the mark field and geometry field from the input sf data.frame.

```{r}
fire_month <- fire_sf %>% 
  select(Month_num)
```

### **6.8.2 Creating ppp**

The code chunk below is used to derive a **ppp** object called *fire_month* from *fire_month* sf data.frame.

```{r}
fire_month_ppp <- as.ppp(fire_month)
fire_month_ppp
```

The code chunk below is used to check the output is in the correct object class.

```{r}
summary(fire_month_ppp)
```

Next, we will check if there are duplicated point events by using the code chunk below.

```{r}
any(duplicated(fire_month_ppp))
```

### **6.8.3 Including Owin object**

The code chunk below is used to combine *origin_am_ppp* and *am_owin* objects into one.

```{r}
fire_month_owin <- fire_month_ppp[kbb_owin]
summary(fire_month_owin)
```

As a good practice, `plot()` is used to plot *ff_owin* so that we can examine the correctness of the output object.

```{r}
plot(fire_month_owin)
```

### **6.8.4 Computing Spatio-temporal KDE**

Next, `spattemp.density()` of sparr package is used to compute the STKDE.

```{r}
library(sparr) # mac need to download and install file first

st_kde <- spattemp.density(fire_month_owin)
summary(st_kde)
```

### **6.8.5 Plotting the spatio-temporal KDE object**

In the code chunk below, `plot()` of R base is used to the KDE for between July 2023 - December 2023.

```{r}
tims <- c(7,8,9,10,11,12)
par(mfcol=c(2,3))
for(i in tims){ 
  plot(st_kde, i, 
       override.par=FALSE, 
       fix.range=TRUE, 
       main=paste("KDE at month",i))
}
```

## **6.9 Computing STKDE by Day of Year**

In this section, you will learn how to computer the STKDE of forest fires by day of year.

### **6.9.1 Creating ppp object**

In the code chunk below, DayofYear field is included in the output ppp object.

```{r}
fire_yday_ppp <- fire_sf %>% 
  select(DayofYear) %>%
  as.ppp()
```

### **6.9.2 Including Owin object**

Next, code chunk below is used to combine the ppp object and the owin object.

```{r}
fire_yday_owin <- fire_yday_ppp[kbb_owin]
summary(fire_yday_owin)
```

### **6.9.3** Spatiotemporal Kernel Density Estimation

```{r}
kde_yday <- spattemp.density(
  fire_yday_owin)
summary(kde_yday)
```

```{r}
plot(kde_yday)
```

## **6.10 Computing STKDE by Day of Year: Improved method**

One of the nice function provides in sparr package is `BOOT.spattemp()`. It support bandwidth selection for standalone spatiotemporal density/intensity based on bootstrap estimation of the MISE, providing an isotropic scalar spatial bandwidth and a scalar temporal bandwidth.

Code chunk below uses `BOOT.spattemp()` to determine both the spatial bandwidth and the scalar temporal bandwidth.

```{r}
set.seed(1234)
BOOT.spattemp(fire_yday_owin) 
```

### **6.10.1 Computing spatio-temporal KDE**

Now, the STKDE will be derived by using h and lambda values derive in previous step.

```{r}
kde_yday <- spattemp.density(fire_yday_owin,
                             h = 9000,
                             lambda = 19)
summary(kde_yday)
```

### **6.10.2 Plotting the output spatio-temporal KDE**

Last, `plot()` of **sparr** package is used to plot the output as shown below.

```{r}
plot(kde_yday)
```
