---
title: "Hands-on Exercise 2"
author: "Jeounghyun Kim"
date-modified: "last-modified" 
execute:
  echo: true
  eval: true
  warning: false
  freeze: true
---

# **2  Thematic Mapping and GeoVisualisation with R**

## **2.1 Overview**

-   **Thematic Mapping**

    -   Uses map symbols to visualize properties of geographic features that are not directly visible.

    -   Examples: population, temperature, crime rate, property prices, etc.

-   **Geovisualisation**

    -   The process of applying visual representations and cartographic techniques to explore, analyze, and communicate geospatial data.

    -   Integrates **cartography, computer science, and information visualization**.

    -   Aims to enhance spatial understanding and support knowledge discovery.

### **2.1.1 Survival Tip**

It is advisable for you to read the functional description of each function before using them.

## **2.2 Getting Started**

In this hands-on exercise, the key R package use is [**tmap**](https://cran.r-project.org/web/packages/tmap/) package in R. Beside **tmap** package, five other R packages will be used. They are:

-   [**readr**](https://readr.tidyverse.org/) for importing delimited text file,

-   [**tidyr**](https://tidyr.tidyverse.org/) for tidying data,

-   [**dplyr**](https://dplyr.tidyverse.org/) for wrangling data and

-   [**sf**](https://cran.r-project.org/web/packages/sf/index.html) for handling geospatial data.

-   [**rvest**](https://rvest.tidyverse.org/) for scraping (or harvesting) data from web pages.

Among the five packages, **readr**, **tidyr** and **dplyr** are part of **tidyverse** package.

The code chunk below will be used to install and load these packages in RStudio.

```{r}
pacman::p_load(sf, tmap, tidyverse, rvest)
```

## 2.3 Importing Data into R

### 2.3.1 The Data

Two datasets will be used to create the choropleth map:

1.  **Master Plan 2019 Subzone Boundary (No Sea) (KML)**

    -   Source: [data.gov.sg](https://data.gov.sg/?utm_source=chatgpt.com)

    -   Type: **Geospatial data**

    -   Description: Contains the geographical boundaries of Singapore at the planning subzone level, based on the URA Master Plan 2019.

2.  **Singapore Residents by Planning Area/Subzone, Age Group, Sex, and Type of Dwelling, June 2024 (CSV)**

    -   File name: `respopagesextod2024.csv`

    -   Source: Department of Statistics, Singapore

    -   Type: **Aspatial data**

    -   Description: Contains demographic and dwelling data. Although it does not include coordinate values, the **PA**(Planning Area) and **SZ** (Subzone) fields serve as unique identifiers to link with the Master Plan 2019 Subzone Boundary data.

### **2.3.2 Importing Geospatial Data into R**

The code chunk below uses the *st_read()* function of **sf** package to import `MP14_SUBZONE_WEB_PL`shapefile into R as a simple feature data frame called `mpsz`.

```{r}
mpsz <- st_read("data/Master_plan_2019/MasterPlan2019SubzoneBoundaryNoSeaKML.kml")
```

#### 2.3.2.1 Tidying data

Function to extract values from the HTML description

```{r}
extract_kml_field <- function(html_text, field_name) {
  if (is.na(html_text) || html_text == "") return(NA_character_)
  
  page <- read_html(html_text)
  rows <- page %>% html_elements("tr")
  
  value <- rows %>%
    keep(~ html_text2(html_element(.x, "th")) == field_name) %>%
    html_element("td") %>%
    html_text2()
  
  if (length(value) == 0) NA_character_ else value
}
```

```{r}
mpsz <- mpsz %>%
  mutate(
    REGION_N = map_chr(Description, extract_kml_field, "REGION_N"),
    PLN_AREA_N = map_chr(Description, extract_kml_field, "PLN_AREA_N"),
    SUBZONE_N = map_chr(Description, extract_kml_field, "SUBZONE_N"),
    SUBZONE_C = map_chr(Description, extract_kml_field, "SUBZONE_C")
  ) %>%
  select(-Name, -Description) %>%
  relocate(geometry, .after = last_col())
```

Let's examine mpsz

```{r}
mpsz
```

Notice that only the first ten records will be displayed. Do you know why?

Yes! This is because R only prints a preview of `sf` objects — by default it shows the first 10 rows to avoid flooding the console.

### **2.3.3 Importing Attribute Data into R**

Next, we will import *respopagesextod2024.csv* file into RStudio and save the file into an tibble dataframe called *popdata*.

The task will be performed by using *read_csv()* function of **readr** package as shown in the code chunk below.

```{r}
popdata <- read_csv("data/respopagesextod2024/respopagesextod2024.csv")
head(popdata)
```

### **2.3.4 Data Preparation**

Before a thematic map can be prepared, you are required to prepare a data table with year 2020 values. The data table should include the variables PA, SZ, YOUNG, ECONOMY ACTIVE, AGED, TOTAL, DEPENDENCY.

-   YOUNG: age group 0 to 4 until age groyup 20 to 24,

-   ECONOMY ACTIVE: age group 25-29 until age group 60-64,

-   AGED: age group 65 and above,

-   TOTAL: all age group, and

-   DEPENDENCY: the ratio between young and aged against economy active group

#### 2.3.4.1 Data wrangling

The following data wrangling and transformation functions will be used:

-   *pivot_wider()* of **tidyr** package, and

-   *mutate()*, *filter()*, *group_by()* and *select()* of **dplyr** package

```{r}
popdata2024 <- popdata %>%
  group_by(PA, SZ, AG) %>%
  summarise(`POP` = sum(`Pop`)) %>%
  ungroup()%>%
  pivot_wider(names_from=AG, 
              values_from=POP) %>%
  mutate(YOUNG = rowSums(.[3:6])
         +rowSums(.[12])) %>%
mutate(`ECONOMY ACTIVE` = rowSums(.[7:11])+
rowSums(.[13:15]))%>%
mutate(`AGED`=rowSums(.[16:21])) %>%
mutate(`TOTAL`=rowSums(.[3:21])) %>%  
mutate(`DEPENDENCY` = (`YOUNG` + `AGED`)
/`ECONOMY ACTIVE`) %>%
  select(`PA`, `SZ`, `YOUNG`, 
       `ECONOMY ACTIVE`, `AGED`, 
       `TOTAL`, `DEPENDENCY`)
```

#### 2.3.4.2 Joining the attribute data and geospatial data

Before we can perform the georelational join, one extra step is required to convert the values in PA and SZ fields to uppercase. This is because the values of PA and SZ fields are made up of upper- and lowercase. On the other, hand the SUBZONE_N and PLN_AREA_N are in uppercase.

```{r}
popdata2024 <- popdata2024 %>%
  mutate_at(.vars = vars(PA, SZ), 
          .funs = list(toupper)) %>%
  filter(`ECONOMY ACTIVE` > 0)
```

Next, *left_join()* of **dplyr** is used to join the geographical data and attribute table using planning subzone name e.g. *SUBZONE_N* and *SZ* as the common identifier.

```{r}
mpsz_pop2024 <- left_join(mpsz, popdata2024,
                          by = c("SUBZONE_N" = "SZ"))
```

Thing to learn from the code chunk above:

-   *left_join()* of **dplyr** package is used with `mpsz` simple feature data frame as the left data table is to ensure that the output will be a simple features data frame.

```{r}
write_rds(mpsz_pop2024, "data/mpsz_pop2024.rds")
```

## **2.4 Choropleth Mapping Geospatial Data Using *tmap***

-   **Definition**: Choropleth mapping visualizes enumeration units (e.g., countries, provinces, states, counties, census units) by filling areas with patterns or graduated colors.

    -   Example: Mapping the spatial distribution of Singapore’s aged population using Master Plan 2014 Subzone boundaries.

-   **Two approaches in `tmap`**:

    1.  **Quick plotting** with `qtm()` → fast and simple thematic maps.

    2.  **Custom plotting** with `tmap` elements → flexible and highly customizable maps.

### **2.4.1 Plotting a choropleth map quickly by using *qtm()***

The easiest and quickest to draw a choropleth map using **tmap** is using *qtm()*. It is concise and provides a good default visualisation in many cases.

The code chunk below will draw a cartographic standard choropleth map as shown below.

```{r}
tmap_mode("plot")
qtm(shp = mpsz_pop2024, 
    fill = "DEPENDENCY")
```

Things to learn from the code chunk above:

-   *tmap_mode()* with “plot” option is used to produce a static map. For interactive mode, “view” option should be used.

-   *fill* argument is used to map the attribute (i.e. DEPENDENCY)

### **2.4.2 Creating a choropleth map by using *tmap*’s elements**

**Limitation of `qtm()`**

-   **Advantage**: `qtm()` is quick and easy for drawing choropleth maps.

-   **Disadvantage**: It provides limited control over the aesthetics of individual layers.

-   **Recommendation**: For high-quality cartographic maps, use **`tmap` drawing elements** instead of only relying on `qtm()`.

```{r}
tm_shape(mpsz_pop2024) +
  tm_polygons(fill = "DEPENDENCY",
              fill.scale = tm_scale_intervals(
                style = "quantile", 
                n = 5,
                values = "brewer.blues"),
              fill.legend = tm_legend(
                title = "Dependency ratio")) +
  tm_title("Distribution of Dependency Ratio by planning subzone") +
  tm_layout(frame = TRUE) +
  tm_borders(fill_alpha = 0.5) +
  tm_compass(type="8star", size = 2) +
  tm_scalebar() +
  tm_grid(alpha =0.2) +
  tm_credits("Source: Planning Sub-zone boundary from Urban Redevelopment Authorithy (URA)\n and Population data from Department of Statistics DOS", 
             position = c("left", "bottom"))
```

In the following sub-section, we will share with you tmap functions that used to plot these elements.

#### 2.4.2.1 Drawing a base map

The basic building block of **tmap** is `tm_shape()` followed by one or more layer elemments such as `tm_polygons()`, `tm_symbols()`, `tm_lines()`, `tm_raster()` and `tm_text()`.

In the code chunk below, `tm_shape()` is used to define the input data (i.e *mpsz_pop2024*) and `tm_polygons()` is used to draw the planning subzone polygons

```{r}
tm_shape(mpsz_pop2024) +
  tm_polygons()
```

::: callout-note
By default, it plots areas of polygons in light gray (gray85) and polygons borders in slightly dark gray (gray25).
:::

#### 2.4.2.2 Drawing a choropleth map using `tm_polygons()`

To draw a choropleth map showing the geographical distribution of a selected variable by planning subzone, we just need to assign the target variable such as *Dependency* to *tm_polygons()*.

```{r}
tm_shape(mpsz_pop2024)+
  tm_polygons(fill = "DEPENDENCY")
```

**Key Points from `tm_polygons()`**

-   **Default classification method**: uses the **“pretty”** interval binning.

-   **Default color scheme**: **`Blues3`** from ColorBrewer.

-   **Missing values**: displayed in **grey** by default.

#### 2.4.2.3 Drawing a choropleth map using *tm_fill()* and \*tm_border()\*\*

Actually, `tm_polygons()` is a wraper of `tm_fill()` and `tm_border()`. `tm_fill()` shades the polygons by using the default colour scheme and `tm_borders()` adds the borders of the polygon features onto the choropleth map.

The code chunk below draws a choropleth map by using `tm_fill()` alone.

```{r}
tm_shape(mpsz_pop2024)+
  tm_fill("DEPENDENCY")
```

Notice that the planning subzones are shared according to the respective dependecy values

To add the boundary of the planning subzones, `tm_borders()` will be used as shown in the code chunk below.

```{r}
tm_shape(mpsz_pop2024)+
  tm_fill("DEPENDENCY") +
  tm_borders()
```

**Notes on `tm_borders()`**

-   **Default**: light-gray border lines are added around polygons.

-   **Transparency**:

    -   Controlled by `fill_alpha` (0 = fully transparent, 1 = opaque).

    -   Default = 1 (usually fully opaque).

-   **Other key arguments**:

    -   `col` → border color

    -   `lwd` → border line width (default = 1)

    -   `lty` → border line type (default = "solid")

```{r}
tm_shape(mpsz_pop2024)+
  tm_fill("DEPENDENCY") +
  tm_borders(col = "grey60",
             lwd = 0.1,
             lty = "dashed")
```

### **2.4.3 Data classification methods of tmap**

#### Data Classification in Choropleth Maps

-   **Purpose**: Reduce many observations into grouped ranges (classes) for easier interpretation.

-   **In `tmap`**: 10 classification methods available:

    -   `fixed`, `sd`, `equal`, `pretty` (default), `quantile`, `kmeans`, `hclust`, `bclust`, `fisher`, `jenks`

-   **Usage**: Specify with the `style` argument in `tm_fill()` or `tm_polygons()`.

#### 2.4.3.1 Plotting choropleth maps with built-in classification methods

The code chunk below shows a quantile data classification that used 5 classes.

```{r}
tm_shape(mpsz_pop2024) +
  tm_polygons(fill = "DEPENDENCY",
              fill.scale = tm_scale_intervals(
                style = "quantile",
                n = 5)) +
  tm_borders(fill_alpha = 0.5)
```

In the code chunk below, *equal* data classification method is used.

```{r}
tm_shape(mpsz_pop2024) +
  tm_polygons(fill = "DEPENDENCY",
              fill.scale = tm_scale_intervals(
                style = "equal",
                n = 5)) +
  tm_borders(fill_alpha = 0.5)
```

Notice that the distribution of quantile data classification method are more evenly distributed then equal data classification method.

> **Warning: Maps Lie!**

> DIY: Using what you had learned, prepare choropleth maps by using different classification methods supported by tmap and compare their differences.
>
> ```{r}
> styles <- c("quantile", "equal", "jenks", "sd") 
> maps_style <- lapply(styles, function(st) {
>   tm_shape(mpsz_pop2024) +
>     tm_polygons("DEPENDENCY",
>                 fill.scale = tm_scale_intervals(style = st, n = 5)) +
>     tm_borders() +
>     tm_layout(title = paste("Style:", st))
> })
> tmap_arrange(tmap = maps_style, ncol = 2)
> ```

> DIY: Preparing choropleth maps by using similar classification method but with different numbers of classes (i.e. 2, 6, 10, 20). Compare the output maps, what observation can you draw?

```{r}
classes <- c(2, 6, 10, 20)
maps_n <- lapply(classes, function(k) {
  tm_shape(mpsz_pop2024) +
    tm_polygons("DEPENDENCY",
                fill.scale = tm_scale_intervals(style = "quantile", n = k)) +
    tm_borders() +
    tm_layout(title = paste("n =", k))
})
tmap_arrange(tmap = maps_n, ncol = 2)

```

#### 2.4.3.2 Plotting choropleth map with custome break

-   **Default behavior**:\
    Built-in classification styles in *tmap* (e.g., `quantile`, `equal`, `jenks`) compute category breaks internally.

-   **Manual override**:\
    To override these defaults, you can explicitly set breakpoints using the `breaks` argument in `tm_scale_intervals()`.

    -   Important: to obtain *n* categories, you must provide *n+1* breakpoints.

    -   Both the minimum and maximum values must be included.

    -   The break values must be in strictly increasing order.

-   **Best practice**:\
    Before defining custom breaks, always check the descriptive statistics of the target variable (e.g., `DEPENDENCY`) to understand its distribution.

    -   Example: `summary(mpsz_pop2024$DEPENDENCY)`

    -   Use measures like min, max, mean, and quartiles to guide meaningful breakpoints.

```{r}
summary(mpsz_pop2024$DEPENDENCY)
```

With reference to the results above, we set break point at 0.60, 0.70, 0.80, and 0.90. In addition, we also need to include a minimum and maximum, which we set at 0 and 100. Our `breaks` vector is thus c(0, 0.60, 0.70, 0.80, 0.90, 1.00)

Now, we will plot the choropleth map by using the code chunk below.

```{r}
tm_shape(mpsz_pop2024)+
  tm_polygons(fill = "DEPENDENCY",
              fill.scale = tm_scale_intervals(
                breaks = c(0, 0.60, 0.70, 0.80, 0.90, 1.00))) +
  tm_borders(fill_alpha = 0.5)
```

### **2.4.4 Colour Scheme**

**tmap** supports colour ramps either defined by the user or a set of predefined colour ramps from the **RColorBrewer** package.

#### 2.4.4.1 Using ColourBrewer palette

To change the colour, we assign the preferred colour to *palette* argument of *values* as shown in the code chunk below.

```{r}
tm_shape(mpsz_pop2024) +
  tm_polygons(fill = "DEPENDENCY",
              fill.scale = tm_scale_intervals(
                style = "quantile",
                n = 5,
                values = "brewer.greens")) +
  tm_borders(fill_alpha = 0.5)
```

Notice that the choropleth map is shaded in green.

To reverse the colour shading, add a “-” prefix.

```{r}
tm_shape(mpsz_pop2024) +
  tm_polygons(fill = "DEPENDENCY",
              fill.scale = tm_scale_intervals(
                style = "quantile",
                n = 5,
                values = "-brewer.greens")) +
  tm_borders(fill_alpha = 0.5)
```

::: callout-note
Notice that the colour scheme has been reversed!
:::

#### 2.4.4.2 Cartographic Furniture

Beside map style, **tmap** also also provides arguments to draw other map furniture such as compass, scale bar and grid lines.

In the code chunk below, `tm_compass()`, `tm_scale_bar()`, `tm_grid()` and `tm_credit()` are used to add compass, scale bar, grid lines and data sources onto the choropleth map.

```{r}
tm_shape(mpsz_pop2024) +
  tm_polygons(fill = "DEPENDENCY",
              fill.scale = tm_scale_intervals(
                style = "quantile",
                n = 5)) +
  tm_borders(fill_alpha = 0.5) +
  tm_compass(type="8star", size = 2) +
  tm_scalebar() +
  tm_grid(lwd = 0.1, alpha = 0.2) +
  tm_credits("Source: data.gov.sg & singstat",
             position = c("left", "bottom"))
```

### **2.4.5 Map Layout**

**Map layout** is the arrangement of all elements in a map (background, frame, typography, scale, aspect ratio, margins, etc.).\
It can be customized in *tmap* with the **`tm_layout()`** function, which controls these common settings.

#### 2.4.5.1 Map Legend

In **tmap**, several *legend* options are provided to change the placement, format and appearance of the legend.

```{r}
tm_shape(mpsz_pop2024) +
  tm_polygons(fill = "DEPENDENCY",
              fill.scale = tm_scale_intervals(
                style = "quantile",
                n = 5),
              fill.legend = tm_legend(
                title = "Dependency ratio")) +
  tm_pos_auto_in() +
  tm_borders(fill_alpha = 0.5) +
  tm_compass(type="8star", size = 2) +
  tm_scalebar() +
  tm_grid(lwd = 0.1, alpha = 0.2) +
  tm_credits("Source: data.gov.sg & singstat",
             position = c("left", "bottom"))
```

#### 2.4.5.2 Map style

**tmap** allows a wide variety of layout settings to be changed. They can be called by using *tmap_style()*.

The code chunk below shows the *classic* style is used.

```{r}
tm_shape(mpsz_pop2024) +
  tm_polygons(fill = "DEPENDENCY",
              fill.scale = tm_scale_intervals(
                style = "quantile",
                n = 5,
                values = "-brewer.greens")) + 
  tm_borders(fill_alpha = 0.5) + 
  tmap_style("natural")
```

To reset the default style, refer to the code chunk below.

```{r}
tmap_style("white")
```

## **2.5 Drawing Small Multiple Choropleth Maps**

**Small multiple maps (facet maps)** show several maps side-by-side or stacked, allowing comparison of spatial patterns across another variable (e.g., time).\
In *tmap*, they can be created in three ways:

1.  Assign multiple values to an aesthetic argument,

2.  Combine stand-alone maps with `tmap_arrange()`,

3.  Use `tm_facets()` with a grouping variable.

### **2.5.1 By assigning multiple values to at least one of the aesthetic arguments**

In this example, small multiple choropleth maps are created by assigning two variables to the visual variable (i.e. fill).

```{r}
tm_shape(mpsz_pop2024) + 
  tm_polygons(
    fill = c("YOUNG", "AGED"),
    fill.legend = 
      tm_legend(position = tm_pos_in(
        "right", "bottom")),
    fill.scale = tm_scale_intervals(
      style = "equal", 
      n = 5,
      values = "brewer.blues")) +
  tm_borders(fill_alpha = 0.5) +
  tmap_style("natural")
```

### **2.5.2 By arrange multiples choropleth maps in a grid layout**

In this example, multiple choropleth maps are created and `tmap_arrnage()` is used to arrnage them in a grid layout.

```{r}
youngmap <- tm_shape(mpsz_pop2024)+ 
  tm_polygons(fill = "YOUNG",
              fill.legend = tm_legend(
                position = tm_pos_in(
                  "right", "bottom"),
                  item.height = 0.8),
              fill.scale = tm_scale_intervals(
                style = "quantile", 
                values = "brewer.blues")) +
  tm_borders(fill_alpha = 0.5) +
  tm_title("Distribution of young population")
                
agedmap <- tm_shape(mpsz_pop2024)+ 
  tm_polygons(fill = "AGED",
              fill.legend = tm_legend(
                position = tm_pos_in(
                  "right", "bottom"),
                item.height = 0.8),
              fill.scale = tm_scale_intervals(
              style = "quantile", 
              values = "brewer.blues")) +
  tm_borders(fill_alpha = 0.5) +
  tm_title("Distribution of aged population")

tmap_arrange(youngmap, agedmap, asp=1, ncol=2)
```

### **2.5.3 By defining a group-by variable in *tm_facets()***

In this example, multiple small choropleth maps are created by using **tm_facets()**.

```{r}
tm_shape(mpsz_pop2024) +
  tm_fill(fill = "DEPENDENCY",
          fill.scale = tm_scale_intervals(
            style = "quantile",
            values = "brewer.blues")) + 
  tm_facets(by = "REGION_N",
            nrow = 2, 
            ncols = 3,
            free.coords=TRUE, 
            drop.units=TRUE) +
  tm_layout(legend.show = TRUE,
            title.position = c("center", "center"), 
            title.size = 20) +
  tm_borders(fill_alpha = 0.5)
```

## **2.6 Mappping Spatial Object Meeting a Selection Criterion**

Instead of creating small multiple choropleth map, you can also use `filter()` of **dplyr** package to select geographical area of interest and plot a choropleth map focus only on the selected region.

```{r}
mpsz_pop2024 %>%
  filter(REGION_N == "CENTRAL REGION") %>%
  tm_shape() +
  tm_polygons(fill = "DEPENDENCY",
              fill.scale = tm_scale_intervals(
                style = "quantile", 
                values = "brewer.greens"),
              fill.legend = tm_legend()) +
  tm_borders(fill_alpha = 0.5)
```

## **2.7 Complementing Thematic Map with Statistical Chart**

**Maps and statistical charts** complement each other: maps highlight spatial patterns, while charts show numerical trends and comparisons. Combining both provides a richer understanding of spatial data.

In *tmap*, charts can be integrated into maps using the **`fill.chart`** argument in map layers and the [**legend char**](https://r-tmap.github.io/tmap/reference/tm_chart.html)**t** feature. (click link)

```{r}
mpsz_pop2024 %>%
  filter(REGION_N == "CENTRAL REGION") %>%
  tm_shape() +
  tm_polygons(fill = "DEPENDENCY",
              fill.scale = tm_scale_intervals(
                style = "quantile", 
                values = "brewer.greens"),
              fill.legend = tm_legend(),
              fill.chart = tm_chart_box()) +
  tm_borders() +
  tm_layout(asp = 0.8)
```

In the code chunk below, We improve the visual representation further by highlighting and lebaling the outliers on the choropleth map.

```{r}
mpsz_selected <- mpsz_pop2024 %>%
  filter(REGION_N == "CENTRAL REGION")

stats <- boxplot.stats(mpsz_selected$DEPENDENCY)

outlier_vals <- stats$out

outlier_sf <- mpsz_selected[mpsz_selected$DEPENDENCY %in% outlier_vals, ]

tm_shape(mpsz_selected) +
  tm_polygons(fill = "DEPENDENCY",
          fill.scale = tm_scale_intervals(
            style = "quantile", 
            values = "brewer.blues"),
          fill.legend = tm_legend(),
          fill.chart = tm_chart_box()) +
  tm_borders(fill_alpha = 0.5) +
tm_shape(outlier_sf) +
  tm_borders(col = "red", lwd = 2) +
  tm_text("SUBZONE_N", col = "red", size = 0.7) +
  tm_layout(asp = 0.8)
```

## **2.8 Creating Interactive Map**

**Interactive maps** allow users to zoom, pan, click for details, and explore overlays, offering a dynamic way to understand spatial data. In *tmap*, you can easily switch between static and interactive views with **`tmap_mode()`**, choosing the mode that best fits your analysis.

By modifying the code chunk in sub-section 2.6, the code chunks below build an interactive map using

```{r}
region_selected <- mpsz_pop2024 %>%
  filter(REGION_N == "CENTRAL REGION")
region_bbox <- st_bbox(region_selected)

stats <- boxplot.stats(region_selected$DEPENDENCY)
outlier_vals <- stats$out
outlier_sf <- region_selected[region_selected$DEPENDENCY %in% outlier_vals, ]

tmap_mode("view")
tm_shape(region_selected, 
         bbox = region_bbox) +
  tm_fill("DEPENDENCY",
          id = "SUBZONE_N",
          popup.vars = c(
            "Name" = "SUBZONE_N", 
            "Dependency" = "DEPENDENCY")) +
  tm_borders() +
  tm_shape(outlier_sf) +
  tm_borders(col = "red", lwd = 2)
```

```{r}
tmap_mode("plot")
```

The interactive map above is far from satisfactory. While we want to encourage users to engage and explore the interactive by zooming in and out of the study area freely. But, users might lost in the cyberspace with too much freedom to zoom-in and zoom-out.

To address this issue, `set_zoom_limits` argument can be used to limit the map extend users can zooming in and out of the map areas as shown below.

```{r}
region_selected <- mpsz_pop2024 %>%
  filter(REGION_N == "CENTRAL REGION")
region_bbox <- st_bbox(region_selected)

stats <- boxplot.stats(region_selected$DEPENDENCY)
outlier_vals <- stats$out
outlier_sf <- region_selected[region_selected$DEPENDENCY %in% outlier_vals, ]

tmap_mode("view")

tm_shape(region_selected, 
         bbox = region_bbox) +
  tm_fill("DEPENDENCY",
          id = "SUBZONE_N",
          popup.vars = c(
            "Name" = "SUBZONE_N", 
            "Dependency" = "DEPENDENCY")) +
  tm_borders() +
  tm_shape(outlier_sf) +
  tm_borders(col = "red", lwd = 2) +
  tm_view(set_zoom_limits = c(12,14))
```

## **2.9 Reference**

<https://r4gdsa.netlify.app/chap02.html>

### **2.9.1 All about tmap package**

-   [tmap: Thematic Maps in R](https://www.jstatsoft.org/article/view/v084i06)

-   [tmap](https://cran.r-project.org/web/packages/tmap/index.html)

-   [tmap: get started!](https://cran.r-project.org/web/packages/tmap/vignettes/tmap-getstarted.html)

-   [tmap: changes in version 2.0](https://cran.r-project.org/web/packages/tmap/vignettes/tmap-changes-v2.html)

-   [tmap: creating thematic maps in a flexible way (useR!2015)](http://von-tijn.nl/tijn/research/presentations/tmap_user2015.pdf)

-   [Exploring and presenting maps with tmap (useR!2017)](http://von-tijn.nl/tijn/research/presentations/tmap_user2017.pdf)

### **2.9.2 Geospatial data wrangling**

-   [sf: Simple Features for R](https://cran.r-project.org/web/packages/sf/index.html)

-   [Simple Features for R: StandardizedSupport for Spatial Vector Data](https://journal.r-project.org/archive/2018/RJ-2018-009/RJ-2018-009.pdf)

-   [Reading, Writing and Converting Simple Features](https://cran.r-project.org/web/packages/sf/vignettes/sf2.html)

### **2.9.3 Data wrangling**

-   [dplyr](https://dplyr.tidyverse.org/)

-   [Tidy data](https://cran.r-project.org/web/packages/tidyr/vignettes/tidy-data.html)

-   [tidyr: Easily Tidy Data with ‘spread()’ and ‘gather()’ Functions](https://cran.r-project.org/web/packages/tidyr/tidyr.pdf)
