---
title: "Hands-on Exercise 2"
author: "Jeounghyun Kim"
date-modified: "last-modified" 
execute:
  echo: true
  eval: true
  warning: false
  freeze: true
---

# **2  Thematic Mapping and GeoVisualisation with R**

## **2.1 Overview**

-   **Thematic Mapping**

    -   Uses map symbols to visualize properties of geographic features that are not directly visible.

    -   Examples: population, temperature, crime rate, property prices, etc.

-   **Geovisualisation**

    -   The process of applying visual representations and cartographic techniques to explore, analyze, and communicate geospatial data.

    -   Integrates **cartography, computer science, and information visualization**.

    -   Aims to enhance spatial understanding and support knowledge discovery.

### **2.1.1 Survival Tip**

It is advisable for you to read the functional description of each function before using them.

## **2.2 Getting Started**

In this hands-on exercise, the key R package use is [**tmap**](https://cran.r-project.org/web/packages/tmap/) package in R. Beside **tmap** package, five other R packages will be used. They are:

-   [**readr**](https://readr.tidyverse.org/) for importing delimited text file,

-   [**tidyr**](https://tidyr.tidyverse.org/) for tidying data,

-   [**dplyr**](https://dplyr.tidyverse.org/) for wrangling data and

-   [**sf**](https://cran.r-project.org/web/packages/sf/index.html) for handling geospatial data.

-   [**rvest**](https://rvest.tidyverse.org/) for scraping (or harvesting) data from web pages.

Among the five packages, **readr**, **tidyr** and **dplyr** are part of **tidyverse** package.

The code chunk below will be used to install and load these packages in RStudio.

```{r}
pacman::p_load(sf, tmap, tidyverse, rvest)
```

### 2.3 Importing Data into R

#### 2.3.1 The Data

Two datasets will be used to create the choropleth map:

1.  **Master Plan 2019 Subzone Boundary (No Sea) (KML)**

    -   Source: [data.gov.sg](https://data.gov.sg/?utm_source=chatgpt.com)

    -   Type: **Geospatial data**

    -   Description: Contains the geographical boundaries of Singapore at the planning subzone level, based on the URA Master Plan 2019.

2.  **Singapore Residents by Planning Area/Subzone, Age Group, Sex, and Type of Dwelling, June 2024 (CSV)**

    -   File name: `respopagesextod2024.csv`

    -   Source: Department of Statistics, Singapore

    -   Type: **Aspatial data**

    -   Description: Contains demographic and dwelling data. Although it does not include coordinate values, the **PA**(Planning Area) and **SZ** (Subzone) fields serve as unique identifiers to link with the Master Plan 2019 Subzone Boundary data.

#### **2.3.2 Importing Geospatial Data into R**

The code chunk below uses the *st_read()* function of **sf** package to import `MP14_SUBZONE_WEB_PL`shapefile into R as a simple feature data frame called `mpsz`.

```{r}
mpsz <- st_read("data/Master_plan_2019/MasterPlan2019SubzoneBoundaryNoSeaKML.kml")
```

##### 2.3.2.1 Tidying data

Function to extract values from the HTML description

```{r}
extract_kml_field <- function(html_text, field_name) {
  if (is.na(html_text) || html_text == "") return(NA_character_)
  
  page <- read_html(html_text)
  rows <- page %>% html_elements("tr")
  
  value <- rows %>%
    keep(~ html_text2(html_element(.x, "th")) == field_name) %>%
    html_element("td") %>%
    html_text2()
  
  if (length(value) == 0) NA_character_ else value
}
```

```{r}
mpsz <- mpsz %>%
  mutate(
    REGION_N = map_chr(Description, extract_kml_field, "REGION_N"),
    PLN_AREA_N = map_chr(Description, extract_kml_field, "PLN_AREA_N"),
    SUBZONE_N = map_chr(Description, extract_kml_field, "SUBZONE_N"),
    SUBZONE_C = map_chr(Description, extract_kml_field, "SUBZONE_C")
  ) %>%
  select(-Name, -Description) %>%
  relocate(geometry, .after = last_col())
```

Let's examine mpsz

```{r}
mpsz
```

Notice that only the first ten records will be displayed. Do you know why?

Yes! This is because R only prints a preview of `sf` objects — by default it shows the first 10 rows to avoid flooding the console.

#### **2.3.3 Importing Attribute Data into R**

Next, we will import *respopagesextod2024.csv* file into RStudio and save the file into an tibble dataframe called *popdata*.

The task will be performed by using *read_csv()* function of **readr** package as shown in the code chunk below.

```{r}
popdata <- read_csv("data/respopagesextod2024/respopagesextod2024.csv")
head(popdata)
```

#### **2.3.4 Data Preparation**

Before a thematic map can be prepared, you are required to prepare a data table with year 2020 values. The data table should include the variables PA, SZ, YOUNG, ECONOMY ACTIVE, AGED, TOTAL, DEPENDENCY.

-   YOUNG: age group 0 to 4 until age groyup 20 to 24,

-   ECONOMY ACTIVE: age group 25-29 until age group 60-64,

-   AGED: age group 65 and above,

-   TOTAL: all age group, and

-   DEPENDENCY: the ratio between young and aged against economy active group

##### 2.3.4.1 Data wrangling

The following data wrangling and transformation functions will be used:

-   *pivot_wider()* of **tidyr** package, and

-   *mutate()*, *filter()*, *group_by()* and *select()* of **dplyr** package

```{r}
popdata2024 <- popdata %>%
  group_by(PA, SZ, AG) %>%
  summarise(`POP` = sum(`Pop`)) %>%
  ungroup()%>%
  pivot_wider(names_from=AG, 
              values_from=POP) %>%
  mutate(YOUNG = rowSums(.[3:6])
         +rowSums(.[12])) %>%
mutate(`ECONOMY ACTIVE` = rowSums(.[7:11])+
rowSums(.[13:15]))%>%
mutate(`AGED`=rowSums(.[16:21])) %>%
mutate(`TOTAL`=rowSums(.[3:21])) %>%  
mutate(`DEPENDENCY` = (`YOUNG` + `AGED`)
/`ECONOMY ACTIVE`) %>%
  select(`PA`, `SZ`, `YOUNG`, 
       `ECONOMY ACTIVE`, `AGED`, 
       `TOTAL`, `DEPENDENCY`)
```

##### 2.3.4.2 Joining the attribute data and geospatial data

Before we can perform the georelational join, one extra step is required to convert the values in PA and SZ fields to uppercase. This is because the values of PA and SZ fields are made up of upper- and lowercase. On the other, hand the SUBZONE_N and PLN_AREA_N are in uppercase.

```{r}
popdata2024 <- popdata2024 %>%
  mutate_at(.vars = vars(PA, SZ), 
          .funs = list(toupper)) %>%
  filter(`ECONOMY ACTIVE` > 0)
```

Next, *left_join()* of **dplyr** is used to join the geographical data and attribute table using planning subzone name e.g. *SUBZONE_N* and *SZ* as the common identifier.

```{r}
mpsz_pop2024 <- left_join(mpsz, popdata2024,
                          by = c("SUBZONE_N" = "SZ"))
```

Thing to learn from the code chunk above:

-   *left_join()* of **dplyr** package is used with `mpsz` simple feature data frame as the left data table is to ensure that the output will be a simple features data frame.

```{r}
write_rds(mpsz_pop2024, "data/mpsz_pop2024.rds")
```
