---
title: "Hands-on Exercise 2"
author: "Jeounghyun Kim"
date-modified: "last-modified" 
execute:
  echo: true
  eval: true
  warning: false
  freeze: true
---

# **2  Thematic Mapping and GeoVisualisation with R**

## **2.1 Overview**

-   **Thematic Mapping**

    -   Uses map symbols to visualize properties of geographic features that are not directly visible.

    -   Examples: population, temperature, crime rate, property prices, etc.

-   **Geovisualisation**

    -   The process of applying visual representations and cartographic techniques to explore, analyze, and communicate geospatial data.

    -   Integrates **cartography, computer science, and information visualization**.

    -   Aims to enhance spatial understanding and support knowledge discovery.

### **2.1.1 Survival Tip**

It is advisable for you to read the functional description of each function before using them.

## **2.2 Getting Started**

In this hands-on exercise, the key R package use is [**tmap**](https://cran.r-project.org/web/packages/tmap/) package in R. Beside **tmap** package, five other R packages will be used. They are:

-   [**readr**](https://readr.tidyverse.org/) for importing delimited text file,

-   [**tidyr**](https://tidyr.tidyverse.org/) for tidying data,

-   [**dplyr**](https://dplyr.tidyverse.org/) for wrangling data and

-   [**sf**](https://cran.r-project.org/web/packages/sf/index.html) for handling geospatial data.

-   [**rvest**](https://rvest.tidyverse.org/) for scraping (or harvesting) data from web pages.

Among the five packages, **readr**, **tidyr** and **dplyr** are part of **tidyverse** package.

The code chunk below will be used to install and load these packages in RStudio.

```{r}
pacman::p_load(sf, tmap, tidyverse, rvest)
```

### 2.3 Importing Data into R

#### 2.3.1 The Data

Two datasets will be used to create the choropleth map:

1.  **Master Plan 2019 Subzone Boundary (No Sea) (KML)**

    -   Source: [data.gov.sg](https://data.gov.sg/?utm_source=chatgpt.com)

    -   Type: **Geospatial data**

    -   Description: Contains the geographical boundaries of Singapore at the planning subzone level, based on the URA Master Plan 2019.

2.  **Singapore Residents by Planning Area/Subzone, Age Group, Sex, and Type of Dwelling, June 2024 (CSV)**

    -   File name: `respopagesextod2024.csv`

    -   Source: Department of Statistics, Singapore

    -   Type: **Aspatial data**

    -   Description: Contains demographic and dwelling data. Although it does not include coordinate values, the **PA**(Planning Area) and **SZ** (Subzone) fields serve as unique identifiers to link with the Master Plan 2019 Subzone Boundary data.

#### **2.3.2 Importing Geospatial Data into R**

The code chunk below uses the *st_read()* function of **sf** package to import `MP14_SUBZONE_WEB_PL`shapefile into R as a simple feature data frame called `mpsz`.

```{r}
mpsz <- st_read("data/Master_plan_2019/MasterPlan2019SubzoneBoundaryNoSeaKML.kml")
```

#### 2.3.2.1 Tidying data

Function to extract values from the HTML description

```{r}
extract_kml_field <- function(html_text, field_name) {
  if (is.na(html_text) || html_text == "") return(NA_character_)
  
  page <- read_html(html_text)
  rows <- page %>% html_elements("tr")
  
  value <- rows %>%
    keep(~ html_text2(html_element(.x, "th")) == field_name) %>%
    html_element("td") %>%
    html_text2()
  
  if (length(value) == 0) NA_character_ else value
}
```

```{r}
mpsz <- mpsz %>%
  mutate(
    REGION_N = map_chr(Description, extract_kml_field, "REGION_N"),
    PLN_AREA_N = map_chr(Description, extract_kml_field, "PLN_AREA_N"),
    SUBZONE_N = map_chr(Description, extract_kml_field, "SUBZONE_N"),
    SUBZONE_C = map_chr(Description, extract_kml_field, "SUBZONE_C")
  ) %>%
  select(-Name, -Description) %>%
  relocate(geometry, .after = last_col())
```

Let's examine mpsz

```{r}
mpsz
```
