---
title: "In-class Exercise 5"
author: "Jeounghyun Kim"
date: "September 25, 2025"
date-modified: "last-modified" 
execute:
  echo: true
  eval: true
  warning: false
  freeze: false
---

```{r}
pacman::p_load(sf, sfdep, tmap, tidyverse)
```

```{r}
hunan <- st_read(dsn = "data/geospatial", 
                 layer = "Hunan")
```

```{r}
hunan2012 <- read_csv("data/aspatial/Hunan_2012.csv")
```

## Combining both data frame by using left join

```{r}
hunan_GDPPC <- left_join(hunan, hunan2012) %>%
  select(1:4, 7, 15)
```

## Global Measures of Spatial Association

### Step 1: Deriving Queen's contiguity weights: sfdep methods

```{r}
wm_q <- hunan_GDPPC %>%
  mutate(nb = st_contiguity(geometry),
         wt = st_weights(nb,
                         style = "W"),
         .before = 1) 
```

### Step 2

```{r}
global_moran_perm(wm_q$GDPPC,
                       wm_q$nb,
                       wm_q$wt,
                  nsim = 99)
```

### Performing Global Moran'I permutation test

```{r}
set.seed(1234)
```

```{r}
global_moran_perm(wm_q$GDPPC,
                       wm_q$nb,
                       wm_q$wt,
                  nsim = 99)
```

### The report

The statistical report on previous tab shows that the p-value is smaller than alpha value of 0.05. Hence, we have enough statistical evidence to reject the null hypothesis that the spatial distribution of GPD per capita are resemble random distribution (i.e. independent from spatial). Because the Moran's I statistics is greater than 0. We can infer that the spatial distribution shows sign of clustering.
